### 2024-10-22 ms_boatnoise_shrimp_goby.Rmd
Changed a lot of the figures and cleaned up the directory storage so that certain folders can be ignored

---
title: "shrimp_goby_boatnoise"
author: "Jack L Manera & Jake M Martin"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    code_download: true
    code_folding: hide
    depth: 4
    number_sections: no
    theme:  cosmo
    toc: yes
    toc_float: yes
    toc_depth: 4
  pdf_document:
    toc: yes
---

```{r setup, include = FALSE}
#kniter setting
knitr::opts_chunk$set(
message = FALSE,
warning = FALSE, # no warnings
cache = TRUE,# Cacheing to save time when kniting
tidy = TRUE
)
```

#1. Data wrangling
##1.1 Loading the required packages
```{r, message=FALSE, results='hide'}
#install.packages("pacman")
pacman::p_load(tidyverse, ggplot2, brms, cmdstanr, broom, broom.mixed, tidybayes, ggthemes, emmeans, ggdist, gridExtra, bayesplot, see, patchwork, ggtext, extraDistr, gghalves, ggpp, rstan)

remotes::install_github("stan-dev/loo")
# Setting colours for vidsalisations
colour1 <- "#97A275"
colour2 <- "#F2B23A"
colour3 <- "#F2711B"
colour4 <- "#8CBEB2"
colour5 <- "#E7C27D"
colour6 <- "#F2A775"
colour11 <- "#666E4F"
colour12 <- "#A38346"
colour13 <- "#9C4811"
```

##1.2 Unblinding the data
```{r}
#assigning the directory so that the code can be universally read on any computer as long as the entire project file is shared
directory <- './2-data/'

#reading in the coded raw BORIS output
BORIS_output <- read.csv("./2-data/BORIS_output_update.csv")

#reading in the decoder
Video_codes <- read.csv("./2-data/Video_codes.csv")

#reading in shrimp species information
shrimp_species <- read.csv("./2-data/species.csv")

#binding the decoder to the BORIS output and revealing the treatment and stage
BORIS_output$code <- str_sub(BORIS_output$video_id, 2,)
BORIS_output$code <- as.factor(BORIS_output$code)
shrimp_species$code <- as.factor(shrimp_species$code)
Video_codes$code <- as.factor(Video_codes$code)
decoded_dataframe <- inner_join(inner_join(BORIS_output, Video_codes, by = c("code")), shrimp_species, by = c("code"))
```

## 1.3 Cross-checking the imported datasets
```{r}
# Find common codes across all data sets
common_codes <- Reduce(intersect, list(BORIS_output$code, shrimp_species$code, Video_codes$code))

# Combine all codes from the three data sets
all_codes <- unique(c(BORIS_output$code, shrimp_species$code, Video_codes$code))

# Find codes not common across all data sets
codes_not_common <- setdiff(all_codes, common_codes)

# Define a function to check which data sets contain each code
check_code_origin <- function(code) {
  origins <- c()

  if(code %in% BORIS_output$code) {
    origins <- c(origins, "BORIS_output")
  }
  if(code %in% shrimp_species$code) {
    origins <- c(origins, "shrimp_species")
  }
  if(code %in% Video_codes$code) {
    origins <- c(origins, "Video_codes")
  }

  return(paste(origins, collapse=", "))
}

# Create a dataframe to store the results
results <- data.frame(code = character(0), origin = character(0), stringsAsFactors = FALSE)

# Loop through each code not common across all data sets
for(code in codes_not_common) {
  origin <- check_code_origin(code)
  results <- rbind(results, data.frame(code = as.character(code), origin = origin, stringsAsFactors = FALSE))
}

# Print the results
print(results)

# Missing observations are identified exclusively within the 'Video_codes' dataframe. This situation suggests that all videos capturing both shrimp and gobies are consistently represented across the datasets. However, there were instances where burrows were recorded and assigned video codes but showed no activity from gobies or shrimp during these recordings. These instances led to the assumption that such burrows were inactive, resulting in their exclusion from the further analysis of the study.
```

##1.4 Cleaning the dataframe
###1.4.1 Column renaming
```{r}
decoded_dataframe <- decoded_dataframe |> 
  rename(shrimp_species = species)

#renaming some of the variables and extracting individual variables from columns that have multiple variables contained in one string
decoded_dataframe <- decoded_dataframe |> 
  dplyr::rename(stage = video_section,
                duration = total_duration_s,
                behaviour = behavior) |>
  dplyr::mutate(date = str_sub(date_file, , 6),
                species = str_sub(subject, 5,),
                duration = as.numeric(duration),
                boat_id = str_sub(date_file, -1, -1),
                video_num = as.numeric(str_sub(video_id, 2)),
                burrow = ceiling(video_num/3),
                individual = str_sub(subject, 1, 1),
                helper = str_sub(date_file, -6,-6),
                stroke = ifelse(helper == "t", 'control', 
                                ifelse((boat_id == "A")|(boat_id == "B")|(boat_id == "C")|(boat_id == "D"), 
                                       '2', '4')))

#creating some unique identifiers to make grouping easier later on for reorganising of the dataframe
decoded_dataframe <- decoded_dataframe |> 
  unite("unique_id", 'burrow','individual', 'species', 'stage', remove = FALSE) |> 
  unite("burrow_species_stage", 'burrow', 'species', 'stage', remove = FALSE) |> 
  unite("burrow_species", 'burrow', 'species', remove = FALSE)
```

###1.4.2 Removing redundant BORIS entries
```{r}
#the next few lines of code are to do with the way BORIS outputs data, where there was redundant empty sections due to both species being independently scored in the software. So here we are filtering out the species specific data from the correlating BORIS species observation and then rebinding the data to remove all of the empty artifact entries.

#extracting just the goby data so that a species variable can be made
goby_dataframe <- decoded_dataframe |> 
  dplyr::filter(species == "Goby") |> 
  dplyr::filter(str_detect(video_id, 'g'))

#extracting just the shrimp data so that a species variable can be made
shrimp_dataframe <- decoded_dataframe |> 
  dplyr::filter(species == "Shrimp")  |>  
  dplyr::filter(str_detect(video_id, 'j'))

#binding the two species specific data frames back together so that there is now no empty 
workingdf1 <- bind_rows(goby_dataframe, shrimp_dataframe)

#filtering for the behviours of interest to this experiment
workingdf2 <- workingdf1 |> 
  group_by(unique_id) |> 
  dplyr::filter(behaviour == 'INSIDE' | behaviour == 'EDGE' | behaviour == 'NEAR' | behaviour == 'AWAY' |behaviour == 'CONTACT' |behaviour == 'APART' ) |>
  mutate(averageduration = sum(duration))

#becuase of the way BORIS outputs data there are redundant individuals in the data, e.g., even if there are only two shrimp in a burrow then BORIS will still output observations for a third shrimp and all value will be 0's. this is what this section of code is removing
workingdf3 <- workingdf2 |> 
  group_by(burrow, species, individual) |> 
  dplyr::filter(sum(averageduration) != 0)
```

###1.4.3 Assigning values to missing entries
```{r}
#now here we are making the assumption that when a goby or shrimp is present in other videos but missing from the video of focus it is either away and out of frame for the entire trial if it is a goby. Otherwise it is inside the burrow the entire time if it is a shrimp. These assumptions are based off of preliminary observation and the general tendencies of each species. Because of this assumption we are assigning the respective value to the behviour in cases where one of the species is known to be present in a burrow but not viewed for the length of the trial. 
fixeddf1 <- workingdf3 |> 
  dplyr::filter(averageduration == 0) |> 
  mutate(duration = ifelse(species == "Goby" & behaviour == "AWAY", 300, duration),
         duration = ifelse(species == "Shrimp" & behaviour == "INSIDE", 300, duration)
         )

fixeddf2 <- workingdf3 |> 
  dplyr::filter(averageduration != 0)

workingdf4 <- rbind(fixeddf2, fixeddf1) |> 
  group_by(burrow_species) |> 
  dplyr::mutate(number_conspecifics = max(as.integer(individual))) |> 
  ungroup()

#now using the number of conspecifics to calculate a duration value that represents the mean time spent doing that activity for that species (i.e. goby or shrimp) in that burrow as we are un able to preserve identity and thus instead must look at species averages for each behaviour. 
workingdf5 <- workingdf4 |> 
  group_by(burrow_species, behaviour, stage) |> 
  dplyr::mutate(mean_duration = mean(duration)) |> 
  ungroup() |> 
  dplyr::filter(individual == 1)

#transforming the data frame into wide format and turning time spent in each zone into a proportion of the total time (i.e. a value between 0 and 1)
wide_data <- workingdf5 |>
  pivot_wider(
    id_cols = c(burrow_species, burrow_species_stage, burrow, stage, species, boat_id, stroke, site, number_conspecifics, shrimp_species),
    names_from = behaviour,
    values_from = mean_duration
  )

wide_data <- wide_data |> 
  mutate(INSIDE = as.numeric(INSIDE),
         EDGE = as.numeric(EDGE),
         NEAR = as.numeric(NEAR),
         AWAY = as.numeric(AWAY),
         CONTACT = as.numeric(CONTACT),
         APART = as.numeric(APART)) |> 
  mutate(total = INSIDE+NEAR+EDGE+AWAY,
         inside = (INSIDE + EDGE)/(INSIDE+EDGE+NEAR+AWAY), 
         away = (NEAR + AWAY)/(INSIDE+EDGE+NEAR+AWAY),
         number_conspecifics = as.factor(number_conspecifics),
         shrimp_contact = CONTACT/(CONTACT+APART)
         ) |> 
  dplyr::select(-INSIDE, -EDGE, -NEAR, -AWAY, -CONTACT, -APART)

#checking if every burrow contains observations for both 'Shrimp' and 'Goby'
burrow_values <- unique(wide_data$burrow) 
has_shrimp <- burrow_values %in% wide_data$burrow[wide_data$species == 'Shrimp'] 
has_goby <- burrow_values %in% wide_data$burrow[wide_data$species == 'Goby'] 
burrows_without_both_species <- burrow_values[!(has_shrimp & has_goby)]
cat("burrows without both 'Shrimp' and 'Goby' species:", paste(burrows_without_both_species, collapse = ', '))

#removing the rows with burrows without both 'Shrimp' and 'Goby' species from the dataframe
wide_data <- wide_data[!wide_data$burrow %in% burrows_without_both_species, ]

wide_data <- wide_data |> 
  mutate(burrow = as.factor(burrow))

#Checking to see the count of observations per burrow
burrow_counts <- table(wide_data$burrow)
print(burrow_counts)

#there are some burrows that are missing stages due to the animal not being observered at all during that observation so we will make some conditional assumptions and fill in the missing values.

#a function to fill the missing stages for each species within each burrow
fill_missing_stages <- function(df) {
  all_stages <- c('Pre', 'During', 'Post')

  missing_stages <- setdiff(all_stages, unique(df$stage))

  if(length(missing_stages) == 0) return(data.frame())

  common_values <- df[1, ]

  new_rows <- data.frame()

  for(stage in missing_stages) {
    new_row <- common_values
    new_row$stage <- stage

    # Set column values based on species
    if(new_row$species == "Goby") {
      new_row$away <- 1
      new_row$inside <- 0
      new_row$total <- 300
      new_row$shrimp_contact <- 0
    } else if(new_row$species == "Shrimp") {
      new_row$away <- 0
      new_row$inside <- 1
      new_row$total <- 300
      new_row$shrimp_contact <- NA
    }

    new_rows <- rbind(new_rows, new_row)
  }

  return(new_rows)
}

#applying the function
fixeddf3 <- wide_data |>
  group_by(burrow_species) |>
  do(fill_missing_stages(.)) |>
  ungroup()

wide_data_filled <- rbind(wide_data, fixeddf3)

#Checking to see the count of observations per burrow
filled_burrow_counts <- table(wide_data_filled$burrow)
print(filled_burrow_counts)
```

###1.4.4 Changing dataframe format 
```{r}
workingdf7 <- wide_data_filled |> 
  mutate(burrow_stage = paste(burrow, stage))

goby_df <- workingdf7 |> 
  dplyr::filter(species == "Goby") |> 
  rename(goby_out = away, goby_total = total, goby_number = number_conspecifics) |> 
  select(-c("inside", "species", "burrow_species", "burrow_species_stage", "shrimp_contact"))

shrimp_df <- workingdf7 |> 
  dplyr::filter(species == "Shrimp") |> 
  rename(shrimp_out = away, shrimp_total = total, shrimp_number = number_conspecifics) |> 
  select(c("shrimp_out", "shrimp_total", "shrimp_number", "burrow_stage", "shrimp_contact"))

final_df <- inner_join(goby_df, shrimp_df, by = "burrow_stage") |> 
  mutate(shrimp_contact = ifelse(is.na(shrimp_contact), ifelse(goby_out == 1, 0, 1 - goby_out), shrimp_contact))

write_csv(final_df, "./2-data/final_df.csv")
```

#2. Statistical analysis using zero-one inflated beta regression
##2.1 Building a global model
```{r}
#building a global model for the proportion of time gobies spent out of the burrow
zoib_formula <- bf(
  # mu (mean) part
  goby_out ~ stroke*stage + goby_number + shrimp_number + shrimp_species + (1 | site/burrow),
  # phi (precision) part
  phi ~ stroke*stage + goby_number + shrimp_number + shrimp_species + (1 | site/burrow),
  # alpha (zero-one-inflation) part
  zoi ~ stroke*stage + goby_number + shrimp_number + shrimp_species + (1 | site/burrow),
  # The one-inflated part, conditional on the 0s
  coi ~ stroke*stage + goby_number + shrimp_number + shrimp_species + (1 | site/burrow),

  family = zero_one_inflated_beta()
)
```

##2.2 Specifying priors
###2.2.1 Checking default priors
```{r}
get_prior(
  zoib_formula,
  data = final_df
)
#we can see that all the beta coefficients are being given a flat prior which means it can take on a value from - infinity to + infinity. All the beta coefficients are on the logit scale so lets plot how changing the value of beta relates to different proportions
```

###2.2.2 Graphing logit~proportion
```{r}
tibble(x = seq(-8, 8, by = 0.1)) |> 
  mutate(y = plogis(x)) |> 
  ggplot(aes(x = x, y = y)) +
  geom_line(size = 1) +
  labs(x = "Logit scale", y = "Probability scale")

#we can see that after -5 or +5 the probability doesn't change much, instead of leaving our priors at the default of + and - infinity, we should say that they are most likely to occur between -5 and +5.
```

###2.2.3 Ploting priors
```{r}
# Generate data points
x <- seq(-20, 20, length.out = 1000)  # Adjusted range for better visualization

# Define parameters for the distributions
loc_logistic <- 0
scale_logistic <- 3

df_student_t <- 3
loc_student_t <- 0
scale_student_t <- 2.5

loc_normal <- 0
scale_normal <- 5

# Calculate PDFs
pdf_logistic <- dlogis(x, location = loc_logistic, scale = scale_logistic)
pdf_student_t <- dt(x, df = df_student_t)
pdf_normal <- dnorm(x, mean = loc_normal, sd = scale_normal)

# Plot the PDFs
df <- data.frame(x = x, Logistic = pdf_logistic, Student_t = pdf_student_t, Normal = pdf_normal)
ggplot(df, aes(x)) +
  geom_line(aes(y = Logistic, color = "Logistic(0, 3)")) + # Default priors of Logistic(0,1) have been broadened as these priors seemed too restrictive
  geom_line(aes(y = Student_t, color = "Student_t(3, 0, 2.5)")) +
  geom_line(aes(y = Normal, color = "Normal(0, 5)")) + #For the beta coefficient prior, we will set it to be normally distributed around 0 with an sd of 5, here is the plotted prior distribution 
  labs(title = "Probability Density Functions",
       x = "x",
       y = "Probability Density") +
  scale_color_manual(name = "Distribution", values = c("Logistic(0, 3)" = "blue", "Student_t(3, 0, 2.5)" = "green", "Normal(0, 5)" = "red")) +
  theme_minimal()
```

###2.2.4 Setting priors 
```{r}
#the priors for the intercept have been left as the default and the coefficient priors defined above
priors <- c(set_prior("student_t(3, 0, 2.5)", class = "Intercept"),
            set_prior("logistic(0, 3)", class = "Intercept", dpar = "coi"),
            set_prior("logistic(0, 3)", class = "Intercept", dpar = "zoi"),
            set_prior("student_t(3, 0, 2.5)", class = "Intercept", dpar = "phi", lb = 0),
            set_prior("student_t(3, 0, 2.5)", class = "sd", lb = 0),
            set_prior("student_t(3, 0, 2.5)", class = "sd", dpar = "phi", lb = 0),
            set_prior("student_t(3, 0, 2.5)", class = "sd", dpar = "zoi", lb = 0),
            set_prior("student_t(3, 0, 2.5)", class = "sd", dpar = "coi", lb = 0),
            set_prior("normal(0, 5)", class = "b"),
            set_prior("normal(0, 5)", class = "b", dpar = "phi"),
            set_prior("normal(0, 5)", class = "b", dpar = "zoi"),
            set_prior("normal(0, 5)", class = "b", dpar = "coi"))
```

###2.2.5 Prior predictive check 
```{r}
prior_formula <- bf(
  # mu (mean) part
  goby_out ~ stroke*stage + goby_number + shrimp_number + shrimp_species + (1 | site/burrow),
  # phi (precision) part
  phi ~ stroke*stage + goby_number + shrimp_number + shrimp_species + (1 | site/burrow),
  # alpha (zero-one-inflation) part
  zoi ~ stroke*stage + goby_number + shrimp_number + shrimp_species + (1 | site/burrow),
  # The one-inflated part, conditional on the 0s
  coi ~ stroke*stage + goby_number + shrimp_number + shrimp_species + (1 | site/burrow),

  family = zero_one_inflated_beta()
)

prior_model <- brm(
  prior_formula,
  data = final_df,
  family = zero_one_inflated_beta(),
  prior = priors,
  control = list(adapt_delta = 0.99,
                 max_treedepth = 12),
  chains = 4, iter = 1000, warmup = 100,
  cores = 4, threads = threading(2),
  backend = "cmdstanr",
  seed = 12345,
  sample_prior = "only",
  file = "prior_check"
)

pp_check(prior_model, ndraws=20)
pp_check(prior_model, ndraws=100, type = 'stat', stat = 'mean')
```

##2.3 Modelling the proportion of time gobies spent outside of the burrow
```{r}
goby_out_formula_1 <- bf(
  # mu (mean) part
  goby_out ~ stroke*stage + goby_number + shrimp_number + shrimp_species + (1 | site/burrow),
  # phi (precision) part
  phi ~ stroke*stage + goby_number + shrimp_number + shrimp_species + (1 | site/burrow),
  # alpha (zero-one-inflation) part
  zoi ~ stroke*stage + goby_number + shrimp_number + shrimp_species + (1 | site/burrow),
  # The one-inflated part, conditional on the 0s
  coi ~ stroke*stage + goby_number + shrimp_number + shrimp_species + (1 | site/burrow),

  family = zero_one_inflated_beta()
)

goby_out_model <- brm(
  goby_out_formula_1,
  data = final_df,
  family = zero_one_inflated_beta(),
  prior = priors,
  control = list(adapt_delta = 0.99,
                 max_treedepth = 12),
  chains = 4, iter = 4000, warmup = 1000,
  cores = 4, threads = threading(2),
  backend = "cmdstanr",
  seed = 12345,
  file = "goby_out_model",

)

summary(goby_out_model)
```
###2.3.1 Model fit and diagnostics
####2.3.1.1 Model convergence
```{r}
#convergence diagnostics (R-hat)
extract_rhats <- function(model) {
  model_summary <- summary(model)
  #extract R-hat values for fixed effects
  fixed_rhats <- model_summary$fixed[, "Rhat", drop = FALSE]
  #check for random effects and extract R-hat values if present
  if (!is.null(model_summary$random)) {
    random_rhats <- do.call(rbind, lapply(model_summary$random, function(x) x[, "Rhat", drop = FALSE]))
    #combine fixed and random effects R-hat values
    rhats <- rbind(fixed_rhats, random_rhats)
  } else {
    rhats <- fixed_rhats
  }

  return(rhats)
}

extract_rhats(goby_out_model)
```

####2.3.1.2 Model fit
```{r}
#posterior predictive checks
pp_check(goby_out_model, ndraws = 50)

pp_check(goby_out_model, type='stat', stat='mean')

#fitted vs predicted values
summary(fitted(goby_out_model))
summary(predict(goby_out_model))
```

###2.3.2 Posterior marginal effects
####2.3.2.1 Pairwise contrasts of stroke and stage
```{r}
goby_out_contrast_effect <- goby_out_model |>
  emmeans(~ stroke*stage, 
          epred = TRUE,
          re_formla = NULL,
          rg.limit =14000) |> 
  contrast(method = "revpairwise")

# Extract the relevant contrast information
goby_out_results.table <- as.data.frame(goby_out_contrast_effect) |> 
  dplyr::filter(contrast %in% c("control Pre - 2 Pre", "control Pre - 4 Pre","4 Pre - 2 Pre", "control Pre - control During", "control Pre - control Post", "4 Pre - 4 During", "4 Pre - 4 Post", "2 Pre - 2 During", "2 Pre - 2 Post"))

goby_out_results.table

goby_out_contrast_draws <- goby_out_model |>
  emmeans(~ stroke*stage, 
          epred = TRUE,
          re_formla = NULL,
          rg.limit =14000) |> 
  contrast(method = "revpairwise") |> 
  gather_emmeans_draws()
  goby_out_contrast_draws

goby_out_temporal_contrasts <- goby_out_contrast_draws |> 
  dplyr::filter(contrast %in% c("control Pre - control During", "control Pre - control Post", "4 Pre - 4 During", "4 Pre - 4 Post", "2 Pre - 2 During", "2 Pre - 2 Post")) |> 
  mutate(stroke = ifelse(contrast %in% c("control Pre - control During", "control Pre - control Post"), "Control", ifelse(contrast %in% c("4 Pre - 4 During", "4 Pre - 4 Post"), "4", "2")))

goby_out_temporal_contrasts$stroke <- factor(goby_out_temporal_contrasts$stroke, levels = c("Control", "4", "2"))

fig.goby_out_temporal_contrasts<- ggplot(goby_out_temporal_contrasts, aes(x = .value, y = factor(contrast, levels = c("control Pre - control During", "control Pre - control Post", "4 Pre - 4 During", "4 Pre - 4 Post", "2 Pre - 2 During", "2 Pre - 2 Post")), color = stroke, fill = stroke)) +
   geom_vline(xintercept = 0, color = "black", size = 0.5) +
  stat_halfeye(.width = c(0.89, 0.95))+
  scale_color_manual(values = c(colour11, colour12, colour13)) +
  scale_fill_manual(values = alpha(c(colour1, colour2, colour3), 0.8)) +
  labs(y = "Contrasts of the time periods relative to noise exposure",
       fill = "Noise treatment", color = "Noise treatment",
       fill_ramp = "Credible interval") +
  xlab(NULL) +
  coord_flip() +
  theme(legend.position = "right", 
        axis.text.x = element_text(angle = 50, hjust =1)) +
  scale_y_discrete(labels = c("Pre - During", "Pre - Post", "Pre - During", "Pre - Post", "Pre - During", "Pre - Post")) +
  scale_x_continuous(breaks = seq(-0.4, 0.5, 0.1), labels = c("-0.4","-0.3", "-0.2","-0.1" , "0", "0.1", "0.2", "0.3", "0.4", "0.5"), limits = c(-0.4, 0.4))

fig.goby_out_temporal_contrasts
```
###2.3.3 Extracting and plotting model estimates
####2.3.3.1 Extracting estimates
```{r}
goby_out_model_estimates <- goby_out_model |> 
  emmeans(~ stroke*stage,
          epred = TRUE,
          rg.limit = 14000,
          re_formula = NA)

goby_out_model_estimates <- as.data.frame(goby_out_model_estimates)

goby_out_model_estimates

write.csv(goby_out_model_estimates, "./2-data/goby_out_model_estimates.csv")

#Storing the model estimate means for each group to plot later on
goby_out_model_estimates_control <- goby_out_model_estimates |> 
  dplyr::filter(stroke == "control") |> 
  rename(.value = emmean)

goby_out_model_estimates_four <- goby_out_model_estimates |> 
  dplyr::filter(stroke == "4")|> 
  rename(.value = emmean)

goby_out_model_estimates_two <- goby_out_model_estimates |> 
  dplyr::filter(stroke == "2")|> 
  rename(.value = emmean)
```
####2.3.3.2 Plotting model estimates and raw data
```{r}
goby_out_model_draws <- goby_out_model |> 
  emmeans(~ stroke*stage,
          epred = TRUE,
          rg.limit = 14000,
          re_formula = NA) |> 
  gather_emmeans_draws()

goby_out_model_draws <- goby_out_model_draws |> 
  rename(emmean = .value)

goby_out_raw_data <- final_df |> 
  rename(emmean = goby_out)

goby_out_raw_data$stage <- factor(goby_out_raw_data$stage, levels = c("Pre", "During", "Post"))
goby_out_raw_data$stroke <- factor(goby_out_raw_data$stroke, levels = c("control", "4", "2"))

goby_out_model_estimates$stage <- factor(goby_out_model_estimates$stage, levels = c("Pre", "During", "Post"))
goby_out_model_estimates$stroke <- factor(goby_out_model_estimates$stroke, levels = c("control", "4", "2"))

goby_out_model_draws$stage <- factor(goby_out_model_draws$stage, levels = c("Pre", "During", "Post"))
goby_out_model_draws$stroke <- factor(goby_out_model_draws$stroke, levels = c("control", "4", "2"))

goby_out_means <- goby_out_raw_data |> 
  group_by(stage, stroke) |> 
  summarise(emmean = mean(emmean))

goby_out_fig_together <- ggplot(data = goby_out_raw_data, aes(x = stage, y = emmean, color = stroke, fill = stroke)) +
  geom_point(data = goby_out_means, alpha = 1, position = position_dodgenudge(width = 0.8, x = -0.09), size = 9, shape = "_") +
  geom_half_point(alpha = 0.3, position = position_dodgenudge(width = 0.8, x = -0.03), side = "l", range_scale = 0.5) +
  geom_half_violin(data = goby_out_model_draws, aes(x = stage, y = emmean, color = stroke, fill = stroke), alpha = 0.6, position = position_dodge(width = 0.8), side = "r", size = 0) +
  geom_point(data = goby_out_model_estimates, aes(x = stage, y = emmean, color = stroke), position = position_dodge(width = 0.8), size = 3) +
  geom_errorbar(data = goby_out_model_estimates, aes(x = stage, y = emmean, ymin = lower.HPD, ymax = upper.HPD, color = stroke), position = position_dodge(width = 0.8), width = 0, size = 1.25) +
  scale_fill_manual(values = c("control" = "#76A34A", "4" = "#F29E00", "2" = "#E25A00"), name = "Noise exposure", guide = guide_legend(override.aes = list(size = 3))) +
  scale_color_manual(values = c("control" = "#2A330E", "4" = "#805213", "2" = "#542709"), name = "Noise exposure", guide = guide_legend(override.aes = list(size = 3))) +
  scale_x_discrete(labels = c("Pre", "During", "Post")) +
  labs(title = "The effects of boat noise on goby burrow use",
       y = "Proportion of time in out of the burrow",
       x = "Experimental phase relative to noise exposure") +
  theme_classic() +
  theme(
    axis.text.x = element_text(size = 10, margin = margin(t = 5)),  
    axis.text.y = element_text(size = 10, margin = margin(r = 5)),  
    axis.title.x = element_text(size = 13, margin = margin(t = 10)),
    axis.title.y = element_text(size = 13, margin = margin(r = 10)),
    legend.text = element_text(size = 10),
    legend.title = element_text(size = 10),
    legend.position = "right",
    legend.justification = "top",
    plot.title = element_text(hjust = 0.5, size = 16),  
    plot.margin = margin(10, 10, 10, 10)
  )

goby_out_fig_together
ggsave("./3-figs/goby_out_fig_together.png", plot = goby_out_fig_together, units = "px", width = 2200, height = 1400, dpi = 300)
```

####2.3.3.3 Plotting control estimates and raw data
```{r}
goby_out_model_draws <- goby_out_model |> 
  emmeans(~ stroke*stage,
          epred = TRUE,
          rg.limit = 14000,
          re_formula = NA) |> 
  gather_emmeans_draws()

goby_out_raw_data <- final_df |>
  rename(.value = goby_out)

goby_out_model_draws$stage <- factor(goby_out_model_draws$stage, levels = c("Pre", "During", "Post"))
goby_out_model_draws$stroke <- factor(goby_out_model_draws$stroke, levels = c("control", "4", "2"))

goby_out_control_model_draws <- goby_out_model_draws |> 
  dplyr::filter(stroke == "control")

goby_control_raw_data <- final_df |> 
  dplyr::filter(stroke =="control") |> 
  rename(.value = goby_out)

goby_control_means <- goby_control_raw_data |> 
  group_by(stage) |> 
  summarise(.value = mean(.value))

goby_control_raw_data$stage <- factor(goby_control_raw_data$stage, levels = c("Pre", "During", "Post"))

#assigning a value for the baseline mean
goby_out_model_estimates_control_intercept <- goby_out_model_estimates_control %>%
  dplyr::filter(stage == "Pre") %>%
  pull(.value)

goby_out_fig_control <- ggplot(goby_control_raw_data, aes(y = .value, x = stage, fill = stage, color = stage)) +
  geom_half_violin(data = goby_out_control_model_draws, alpha = 0.7, size = 0, side = "r", position = position_nudge(x = 0.2)) +
  geom_jitter(position = position_jitternudge(x = -0.1, width = 0.1, nudge.from = "jittered"), alpha = 0.4, size = 2) +
  geom_crossbar(alpha = 0.6, size = 0.8, data = goby_out_model_estimates_control, aes(ymin = lower.HPD, ymax = upper.HPD), width = 0, position = position_nudge(x = 0.2))+
  geom_point(pch = 23, stroke = 1.2, data = goby_out_model_estimates_control, size = 2, position = position_nudge(x = 0.2)) +
  geom_point(data = goby_control_means, position = position_dodgenudge(width = 0.8, x = -0.1), size = 9, shape = "_") +
  labs(x="Period relative to treatment",
       y="Proportion of time out of burrow") +
  theme(legend.position = "none", 
        axis.title.x = element_blank(), 
        axis.text.x = element_text(size = 13),  
        axis.text.y = element_text(size = 13),
        axis.title.y = element_text(size = 16, face = "bold", margin = margin(t = 0, b = 0, l = 0, r = 10)),
        plot.title = element_textbox_simple(face = "bold", size = 16, halign = 0.5, linewidth = 0.3, linetype = 1, color = "black", fill = "#A3C088", padding = margin(5, 5, 5, 5)), 
        panel.background = element_rect(fill = "white", colour = "black"),
        plot.background = element_rect(fill = NA, colour = NA))+
  ggtitle("Control")+
  scale_fill_manual(values = c("Pre" = "darkgrey", "During" = "#76A34A", "Post" = "#D3D9A7")) +
  scale_color_manual(values = c("Pre" = "#242424", "During" = "#2A330E", "Post" = "#5d634c"))

goby_out_fig_control
```
####2.3.3.4 Plotting 4-stroke estimates and raw data
```{r}
goby_out_control_model_draws <- goby_out_model_draws |> 
  dplyr::filter(stroke == "4")

goby_four_raw_data <- final_df |> 
  dplyr::filter(stroke =="4") |> 
  rename(.value = goby_out)

goby_four_raw_data$stage <- factor(goby_four_raw_data$stage, levels = c("Pre", "During", "Post"))

goby_four_means <- goby_four_raw_data |> 
  group_by(stage) |> 
  summarise(.value = mean(.value))

goby_out_fig_four <- ggplot(goby_four_raw_data, aes(y = .value, x = stage, fill = stage, color = stage)) +
  geom_half_violin(data = goby_out_control_model_draws, alpha = 0.7, size = 0, side = "r", position = position_nudge(x = 0.2)) +
  geom_jitter(position = position_jitternudge(x = -0.1, width = 0.1, nudge.from = "jittered"), alpha = 0.4, size = 2) +
  geom_crossbar(alpha = 0.6, size = 0.8, data = goby_out_model_estimates_four, aes(ymin = lower.HPD, ymax = upper.HPD), width = 0, position = position_nudge(x = 0.2))+
  geom_point(pch = 23, stroke = 1.2, data = goby_out_model_estimates_four, size = 2, position = position_nudge(x = 0.2)) +
  geom_point(data = goby_four_means, position = position_dodgenudge(width = 0.8, x = -0.1), size = 9, shape = "_") +
  xlab("Experimental phase relative to noise exposure")+
  theme(legend.position = "none", 
        axis.title.y = element_blank(),  
        axis.text.x = element_text(size = 13),  
        axis.text.y = element_blank(), 
        axis.ticks.y = element_blank(), 
        axis.title.x = element_text(size = 16, face = "bold", margin = margin(t = 15, b = 0, l = 0, r = 0)),
        plot.title = element_textbox_simple(face = "bold", size = 16, halign = 0.5, linewidth = 0.3, linetype = 1, color = "black", fill = "#F6BC65", padding = margin(5, 5, 5, 5)), 
        panel.background = element_rect(fill = "white", colour = "black"),
        plot.background = element_rect(fill = NA, colour = NA))+
  ggtitle("4-stroke")+
  scale_fill_manual(values = c("Pre" = "darkgrey", "During" = "#F29E00", "Post" = "#F2D5A0")) +
  scale_color_manual(values = c("Pre" = "#242424", "During" = "#805213", "Post" = "#70532b"))

goby_out_fig_four
```

####2.3.3.5 Plotting 2-stroke estimates and raw data
```{r}
goby_out_control_model_draws <- goby_out_model_draws |> 
  dplyr::filter(stroke == "2")

goby_two_raw_data <- final_df |> 
  dplyr::filter(stroke =="2") |> 
  rename(.value = goby_out)

goby_two_raw_data$stage <- factor(goby_two_raw_data$stage, levels = c("Pre", "During", "Post"))

goby_two_means <- goby_two_raw_data |> 
  group_by(stage) |> 
  summarise(.value = mean(.value))

goby_out_fig_two <- ggplot(goby_two_raw_data, aes(y = .value, x = stage, fill = stage, color = stage)) +
  geom_half_violin(data = goby_out_control_model_draws, alpha = 0.7, size = 0, side = "r", position = position_nudge(x = 0.2)) +
  geom_jitter(position = position_jitternudge(x = -0.1, width = 0.1, nudge.from = "jittered"), alpha = 0.4, size = 2) +
  geom_crossbar(alpha = 0.6, size = 0.8, data = goby_out_model_estimates_two, aes(ymin = lower.HPD, ymax = upper.HPD), width = 0, position = position_nudge(x = 0.2))+
  geom_point(pch = 23, stroke = 1.2, data = goby_out_model_estimates_two, size = 2, position = position_nudge(x = 0.2)) +
  geom_point(data = goby_two_means, position = position_dodgenudge(width = 0.8, x = -0.1), size = 9, shape = "_") +
  theme(legend.position = "none", 
        axis.title.x = element_blank(),  
        axis.text.x = element_text(size = 13),
        axis.title.y = element_blank(), 
        axis.text.y = element_blank(), 
        axis.ticks.y = element_blank(), 
        plot.title = element_textbox_simple(face = "bold", size = 16, halign = 0.5, linewidth = 0.3, linetype = 1, color = "black", fill = "#EB9063", padding = margin(5, 5, 5, 5)), 
        panel.background = element_rect(fill = "white", colour = "black"),
        plot.background = element_rect(fill = NA, colour = NA))+
  ggtitle("2-stroke") +
  scale_fill_manual(values = c("Pre" = "darkgrey", "During" = "#E25A00", "Post" = "#F2B389")) +
  scale_color_manual(values = c("Pre" = "#242424", "During" = "#542709", "Post" = "#66422a"))

goby_out_fig_two
```
####2.3.3.6 Combining plots into one figure
```{r}
goby_out_plot3 <- goby_out_fig_control + goby_out_fig_four + goby_out_fig_two + 
  plot_layout(ncol = 3)

goby_out_plot3
ggsave("./3-figs/goby_out_plot3.png", plot = goby_out_plot3, units = "px", width = 2200, height = 1400, dpi = 300)
```

###2.3.4 Effects of covariates on the proportion of time gobies spent outside of the burrow
####2.3.4.1 Goby number
```{r}
goby_out_goby_number <- goby_out_model |>
  emmeans(~ goby_number, 
          epred = TRUE,
          re_formla = NULL) |> 
  contrast(method = "revpairwise")

goby_out_goby_number

goby_out_goby_number_draws <- goby_out_model |>
  emmeans(~ goby_number, 
          epred = TRUE,
          re_formla = NULL) |> 
  contrast(method = "revpairwise")|> 
  gather_emmeans_draws()

goby_out_goby_number_fig<- ggplot(goby_out_goby_number_draws, aes(x = .value, y = contrast)) +
   geom_vline(xintercept = 0, color = "black", size = 0.5) +
  stat_halfeye(.width = c(0.95))+
  labs(y = "Contrasts of the effect of the number of goby",
       fill = "Noise treatment", color = "Noise treatment",
       fill_ramp = "Credible interval") +
  xlab(NULL) +
  coord_flip()

goby_out_goby_number_fig
```

####2.3.4.2 Shrimp number
```{r}
goby_out_shrimp_number <- goby_out_model |>
  emmeans(~ shrimp_number, 
          epred = TRUE,
          re_formla = NULL) |> 
  contrast(method = "revpairwise")

goby_out_shrimp_number

goby_out_shrimp_number_estimates <- goby_out_model |>
  emmeans(~ shrimp_number, 
          epred = TRUE,
          re_formla = NULL)

goby_out_shrimp_number_estimates

goby_out_shrimp_number_draws <- goby_out_model |>
  emmeans(~ shrimp_number, 
          epred = TRUE,
          re_formla = NULL) |> 
  contrast(method = "revpairwise")|> 
  gather_emmeans_draws()

goby_out_shrimp_number_fig<- ggplot(goby_out_shrimp_number_draws, aes(x = .value, y = contrast)) +
   geom_vline(xintercept = 0, color = "black", size = 0.5) +
  stat_halfeye(.width = c(0.95))+
  labs(y = "Contrasts of the effect of the number of shrimp",
       fill = "Noise treatment", color = "Noise treatment",
       fill_ramp = "Credible interval") +
  xlab(NULL) +
  coord_flip()

goby_out_shrimp_number_fig
```

####2.3.4.3 Shrimp species
```{r}
goby_out_shrimp_species <- goby_out_model |>
  emmeans(~ shrimp_species, 
          epred = TRUE,
          re_formla = NULL) |> 
  contrast(method = "revpairwise")

goby_out_shrimp_species

goby_out_shrimp_species_estimates <- goby_out_model |>
  emmeans(~ shrimp_species, 
          epred = TRUE,
          re_formla = NULL)

goby_out_shrimp_species_estimates

goby_out_shrimp_species_draws <- goby_out_model |>
  emmeans(~ shrimp_species, 
          epred = TRUE,
          re_formla = NULL) |> 
  contrast(method = "revpairwise")|> 
  gather_emmeans_draws()

goby_out_shrimp_species_fig<- ggplot(goby_out_shrimp_species_draws, aes(x = .value, y = contrast)) +
   geom_vline(xintercept = 0, color = "black", size = 0.5) +
  stat_halfeye(.width = c(0.95))+
  labs(y = "Contrasts of the effect of the species of shrimp",
       fill = "Noise treatment", color = "Noise treatment",
       fill_ramp = "Credible interval") +
  xlab(NULL) +
  coord_flip()

goby_out_shrimp_species_fig
```

##2.4 Modelling the proportion of time shrimp spent outside of the burrow
```{r}
shrimp_out_formula <- bf(
  # mu (mean) part
  shrimp_out ~ stroke*stage + shrimp_number + goby_number + shrimp_species + (1|site/burrow),
  # phi (precision) part
  phi ~ stroke*stage + shrimp_number + goby_number + shrimp_species + (1|site/burrow),
  # alpha (zero-one-inflation) part
  zoi ~ stroke*stage + shrimp_number + goby_number + shrimp_species + (1|site/burrow),
  # The one-inflated part, conditional on the 0s
  coi ~ stroke*stage + shrimp_number + goby_number + shrimp_species + (1|site/burrow),

  family = zero_one_inflated_beta()
)

shrimp_out_model <- brm(
  shrimp_out_formula,
  data = final_df,
  family = zero_one_inflated_beta(),
  prior = priors,
  control = list(adapt_delta = 0.99,
                 max_treedepth = 12),
  chains = 4, iter = 4000, warmup = 1000,
  cores = 4, threads = threading(2),
  backend = "cmdstanr",
  seed = 12345,
  file = "shrimp_out_model"
)

summary(shrimp_out_model)
```

###2.4.1 Model fit and diagnostics
####2.4.1.1 Model convergence
```{r}
extract_rhats(shrimp_out_model)
```

####2.4.1.2 Model fit
```{r}
#posterior predictive checks
pp_check(shrimp_out_model)

pp_check(shrimp_out_model, type='stat', stat='mean')

#fitted vs predicted values
summary(fitted(shrimp_out_model))
summary(predict(shrimp_out_model))
```

###2.4.2 Posterior marginal effects
####2.4.2.1 Pairwise contrasts of stroke and stage
```{r}
shrimp_out_contrasts <- shrimp_out_model |>
  emmeans(~ stroke*stage, 
          epred = TRUE,
          re_formla = NULL,
          rg.limit =14000) |> 
  contrast(method = "revpairwise")

# Extract the relevant contrast information
shrimp_out_contrasts <- as.data.frame(shrimp_out_contrasts) |> 
  dplyr::filter(contrast %in% c("control Pre - 2 Pre", "control Pre - 4 Pre","4 Pre - 2 Pre", "control Pre - control During", "control Pre - control Post", "4 Pre - 4 During", "4 Pre - 4 Post", "2 Pre - 2 During", "2 Pre - 2 Post"))

shrimp_out_contrasts

shrimp_out_contrast_draws <- shrimp_out_model  |>
  emmeans(~ stroke*stage, 
          epred = TRUE,
          re_formula = NA) |> 
  contrast(method = "revpairwise") |> 
  gather_emmeans_draws()

shrimp_out_temporal_contrasts <- shrimp_out_contrast_draws |> 
  dplyr::filter(contrast %in% c("control Pre - control During", "control Pre - control Post", "4 Pre - 4 During", "4 Pre - 4 Post", "2 Pre - 2 During", "2 Pre - 2 Post")) |> 
  mutate(stroke = ifelse(contrast %in% c("control Pre - control During", "control Pre - control Post"), "Control", ifelse(contrast %in% c("4 Pre - 4 During", "4 Pre - 4 Post"), "4", "2")))

shrimp_out_temporal_contrasts$stroke <- factor(shrimp_out_temporal_contrasts$stroke, levels = c("Control", "4", "2"))

shrimp_out_temporal_contrasts_fig <- ggplot(shrimp_out_temporal_contrasts, aes(x = .value, y = factor(contrast, levels = c("control Pre - control During", "control Pre - control Post", "4 Pre - 4 During", "4 Pre - 4 Post",  "2 Pre - 2 During", "2 Pre - 2 Post")), color = stroke, fill = stroke)) +
  geom_vline(xintercept = 0, color = "black", size = 0.5) +
  stat_halfeye(.width = c(0.89, 0.95)) +
  scale_color_manual(values = c(colour11, colour12, colour13)) +
  scale_fill_manual(values = alpha(c(colour1, colour2, colour3), 0.8)) +
  labs(x = "Difference in the proportion of time spent out of the burrow",
       y = "Contrasts of the time periods relative to noise exposure",
       fill = "Noise treatment", color = "Noise treatment",
       fill_ramp = "Credible interval") +
  coord_flip() +
  theme(legend.position = "top", 
        axis.text.x = element_text(angle = 50, hjust =1)) +
  scale_y_discrete(labels = c("Pre - During", "Pre - Post", "Pre - During", "Pre - Post", "Pre - During", "Pre - Post")) +
  scale_x_continuous(breaks = seq(-0.2, 0.3, 0.1), labels = c("-0.2","-0.1" , "0", "0.1", "0.2", "0.3"), limits = c(-0.3, 0.4))

shrimp_out_temporal_contrasts_fig
```

###2.4.3 Extracting and plotting model estimates
####2.4.3.1 Extracting estimates
```{r}
shrimp_out_model_estimates <- shrimp_out_model |> 
  emmeans(~ stroke*stage,
          epred = TRUE,
          rg.limit = 14000,
          re_formula = NA)

shrimp_out_model_estimates <- as.data.frame(shrimp_out_model_estimates)

shrimp_out_model_estimates

write.csv(shrimp_out_model_estimates, "./2-data/shrimp_out_model_estimates.csv")

#Storing the model estimate means for each group to plot later on
shrimp_out_model_estimates_control <- shrimp_out_model_estimates |> 
  dplyr::filter(stroke == "control") |> 
  rename(.value = emmean)

shrimp_out_model_estimates_four <- shrimp_out_model_estimates |> 
  dplyr::filter(stroke == "4")|> 
  rename(.value = emmean)

shrimp_out_model_estimates_two <- shrimp_out_model_estimates |> 
  dplyr::filter(stroke == "2")|> 
  rename(.value = emmean)
```
####2.4.3.2 Plotting model estimates and raw data
```{r}
shrimp_out_model_draws <- shrimp_out_model |> 
  emmeans(~ stroke*stage,
          epred = TRUE,
          rg.limit = 14000,
          re_formula = NA) |> 
  gather_emmeans_draws()

shrimp_out_model_draws <- shrimp_out_model_draws |> 
  rename(emmean = .value)

shrimp_out_raw_data <- final_df |> 
  rename(emmean = shrimp_out)

shrimp_out_raw_data$stage <- factor(shrimp_out_raw_data$stage, levels = c("Pre", "During", "Post"))
shrimp_out_raw_data$stroke <- factor(shrimp_out_raw_data$stroke, levels = c("control", "4", "2"))

shrimp_out_model_estimates$stage <- factor(shrimp_out_model_estimates$stage, levels = c("Pre", "During", "Post"))
shrimp_out_model_estimates$stroke <- factor(shrimp_out_model_estimates$stroke, levels = c("control", "4", "2"))

shrimp_out_model_draws$stage <- factor(shrimp_out_model_draws$stage, levels = c("Pre", "During", "Post"))
shrimp_out_model_draws$stroke <- factor(shrimp_out_model_draws$stroke, levels = c("control", "4", "2"))

shrimp_out_means <- shrimp_out_raw_data |> 
  group_by(stage, stroke) |> 
  summarise(emmean = mean(emmean))

shrimp_out_fig_together <- ggplot(data = shrimp_out_raw_data, aes(x = stage, y = emmean, color = stroke, fill = stroke)) +
  geom_point(data = shrimp_out_means, alpha = 1, position = position_dodgenudge(width = 0.8, x = -0.09), size = 9, shape = "_") +
  geom_half_point(alpha = 0.3, position = position_dodgenudge(width = 0.8, x = -0.03), side = "l", range_scale = 0.5) +
  geom_half_violin(data = shrimp_out_model_draws, aes(x = stage, y = emmean, color = stroke, fill = stroke), alpha = 0.6, position = position_dodge(width = 0.8), side = "r", size = 0) +
  geom_point(data = shrimp_out_model_estimates, aes(x = stage, y = emmean, color = stroke), position = position_dodge(width = 0.8), size = 3) +
  geom_errorbar(data = shrimp_out_model_estimates, aes(x = stage, y = emmean, ymin = lower.HPD, ymax = upper.HPD, color = stroke), position = position_dodge(width = 0.8), width = 0, size = 1.25) +
  scale_fill_manual(values = c("control" = "#76A34A", "4" = "#F29E00", "2" = "#E25A00"), name = "Noise exposure", guide = guide_legend(override.aes = list(size = 3))) +
  scale_color_manual(values = c("control" = "#2A330E", "4" = "#805213", "2" = "#542709"), name = "Noise exposure", guide = guide_legend(override.aes = list(size = 3))) +
  scale_x_discrete(labels = c("Pre", "During", "Post")) +
  labs(title = "The effects of boat noise on shrimp burrow use",
       y = "Proportion of time in out of the burrow",
       x = "Experimental phase relative to noise exposure") +
  theme_classic() +
  theme(
    axis.text.x = element_text(size = 10, margin = margin(t = 5)),  
    axis.text.y = element_text(size = 10, margin = margin(r = 5)),  
    axis.title.x = element_text(size = 13, margin = margin(t = 10)),
    axis.title.y = element_text(size = 13, margin = margin(r = 10)),
    legend.text = element_text(size = 10),
    legend.title = element_text(size = 10),
    legend.position = "right",
    legend.justification = "top",
    plot.title = element_text(hjust = 0.5, size = 16),  
    plot.margin = margin(10, 10, 10, 10)
  )

shrimp_out_fig_together
ggsave("./3-figs/shrimp_out_fig_together.png", plot = shrimp_out_fig_together, units = "px", width = 2200, height = 1400, dpi = 300)
```

####2.4.3.3 Plotting control estimates and raw data
```{r}
shrimp_out_model_draws <- shrimp_out_model |> 
  emmeans(~ stroke*stage,
          epred = TRUE,
          rg.limit = 14000,
          re_formula = NA) |> 
  gather_emmeans_draws()

shrimp_out_raw_data <- final_df |>
  rename(.value = shrimp_out)

shrimp_out_model_draws$stage <- factor(shrimp_out_model_draws$stage, levels = c("Pre", "During", "Post"))
shrimp_out_model_draws$stroke <- factor(shrimp_out_model_draws$stroke, levels = c("control", "4", "2"))

shrimp_out_control_model_draws <- shrimp_out_model_draws |> 
  dplyr::filter(stroke == "control")

shrimp_control_raw_data <- final_df |> 
  dplyr::filter(stroke =="control") |> 
  rename(.value = shrimp_out)

shrimp_control_means <- shrimp_control_raw_data |> 
  group_by(stage) |> 
  summarise(.value = mean(.value))

shrimp_control_raw_data$stage <- factor(shrimp_control_raw_data$stage, levels = c("Pre", "During", "Post"))

#assigning a value for the baseline mean
shrimp_out_model_estimates_control_intercept <- shrimp_out_model_estimates_control %>%
  dplyr::filter(stage == "Pre") %>%
  pull(.value)

shrimp_out_fig_control <- ggplot(shrimp_control_raw_data, aes(y = .value, x = stage, fill = stage, color = stage)) +
  geom_half_violin(data = shrimp_out_control_model_draws, alpha = 0.7, size = 0, side = "r", position = position_nudge(x = 0.2)) +
  geom_jitter(position = position_jitternudge(x = -0.1, width = 0.1, nudge.from = "jittered"), alpha = 0.4, size = 2) +
  geom_crossbar(alpha = 0.6, size = 0.8, data = shrimp_out_model_estimates_control, aes(ymin = lower.HPD, ymax = upper.HPD), width = 0, position = position_nudge(x = 0.2))+
  geom_point(pch = 23, stroke = 1.2, data = shrimp_out_model_estimates_control, size = 2, position = position_nudge(x = 0.2)) +
  geom_point(data = shrimp_control_means, position = position_dodgenudge(width = 0.8, x = -0.1), size = 9, shape = "_") +
  labs(x="Period relative to treatment",
       y="Proportion of time out of burrow") +
  theme(legend.position = "none", 
        axis.title.x = element_blank(), 
        axis.text.x = element_text(size = 13),  
        axis.text.y = element_text(size = 13),
        axis.title.y = element_text(size = 16, face = "bold", margin = margin(t = 0, b = 0, l = 0, r = 10)),
        plot.title = element_textbox_simple(face = "bold", size = 16, halign = 0.5, linewidth = 0.3, linetype = 1, color = "black", fill = "#A3C088", padding = margin(5, 5, 5, 5)), 
        panel.background = element_rect(fill = "white", colour = "black"),
        plot.background = element_rect(fill = NA, colour = NA))+
  ggtitle("Control")+
  scale_fill_manual(values = c("Pre" = "darkgrey", "During" = "#76A34A", "Post" = "#D3D9A7")) +
  scale_color_manual(values = c("Pre" = "#242424", "During" = "#2A330E", "Post" = "#5d634c"))

shrimp_out_fig_control
```
####2.4.3.4 Plotting 4-stroke estimates and raw data
```{r}
shrimp_out_control_model_draws <- shrimp_out_model_draws |> 
  dplyr::filter(stroke == "4")

shrimp_four_raw_data <- final_df |> 
  dplyr::filter(stroke =="4") |> 
  rename(.value = shrimp_out)

shrimp_four_raw_data$stage <- factor(shrimp_four_raw_data$stage, levels = c("Pre", "During", "Post"))

shrimp_four_means <- shrimp_four_raw_data |> 
  group_by(stage) |> 
  summarise(.value = mean(.value))

shrimp_out_fig_four <- ggplot(shrimp_four_raw_data, aes(y = .value, x = stage, fill = stage, color = stage)) +
  geom_half_violin(data = shrimp_out_control_model_draws, alpha = 0.7, size = 0, side = "r", position = position_nudge(x = 0.2)) +
  geom_jitter(position = position_jitternudge(x = -0.1, width = 0.1, nudge.from = "jittered"), alpha = 0.4, size = 2) +
  geom_crossbar(alpha = 0.6, size = 0.8, data = shrimp_out_model_estimates_four, aes(ymin = lower.HPD, ymax = upper.HPD), width = 0, position = position_nudge(x = 0.2))+
  geom_point(pch = 23, stroke = 1.2, data = shrimp_out_model_estimates_four, size = 2, position = position_nudge(x = 0.2)) +
  geom_point(data = shrimp_four_means, position = position_dodgenudge(width = 0.8, x = -0.1), size = 9, shape = "_") +
  xlab("Experimental phase relative to noise exposure")+
  theme(legend.position = "none", 
        axis.title.y = element_blank(),  
        axis.text.x = element_text(size = 13),  
        axis.text.y = element_blank(), 
        axis.ticks.y = element_blank(), 
        axis.title.x = element_text(size = 16, face = "bold", margin = margin(t = 15, b = 0, l = 0, r = 0)),
        plot.title = element_textbox_simple(face = "bold", size = 16, halign = 0.5, linewidth = 0.3, linetype = 1, color = "black", fill = "#F6BC65", padding = margin(5, 5, 5, 5)), 
        panel.background = element_rect(fill = "white", colour = "black"),
        plot.background = element_rect(fill = NA, colour = NA))+
  ggtitle("4-stroke")+
  scale_fill_manual(values = c("Pre" = "darkgrey", "During" = "#F29E00", "Post" = "#F2D5A0")) +
  scale_color_manual(values = c("Pre" = "#242424", "During" = "#805213", "Post" = "#70532b"))

shrimp_out_fig_four
```

####2.4.3.5 Plotting 2-stroke estimates and raw data
```{r}
shrimp_out_control_model_draws <- shrimp_out_model_draws |> 
  dplyr::filter(stroke == "2")

shrimp_two_raw_data <- final_df |> 
  dplyr::filter(stroke =="2") |> 
  rename(.value = shrimp_out)

shrimp_two_raw_data$stage <- factor(shrimp_two_raw_data$stage, levels = c("Pre", "During", "Post"))

shrimp_two_means <- shrimp_two_raw_data |> 
  group_by(stage) |> 
  summarise(.value = mean(.value))

shrimp_out_fig_two <- ggplot(shrimp_two_raw_data, aes(y = .value, x = stage, fill = stage, color = stage)) +
  geom_half_violin(data = shrimp_out_control_model_draws, alpha = 0.7, size = 0, side = "r", position = position_nudge(x = 0.2)) +
  geom_jitter(position = position_jitternudge(x = -0.1, width = 0.1, nudge.from = "jittered"), alpha = 0.4, size = 2) +
  geom_crossbar(alpha = 0.6, size = 0.8, data = shrimp_out_model_estimates_two, aes(ymin = lower.HPD, ymax = upper.HPD), width = 0, position = position_nudge(x = 0.2))+
  geom_point(pch = 23, stroke = 1.2, data = shrimp_out_model_estimates_two, size = 2, position = position_nudge(x = 0.2)) +
  geom_point(data = shrimp_two_means, position = position_dodgenudge(width = 0.8, x = -0.1), size = 9, shape = "_") +
  theme(legend.position = "none", 
        axis.title.x = element_blank(),  
        axis.text.x = element_text(size = 13),
        axis.title.y = element_blank(), 
        axis.text.y = element_blank(), 
        axis.ticks.y = element_blank(), 
        plot.title = element_textbox_simple(face = "bold", size = 16, halign = 0.5, linewidth = 0.3, linetype = 1, color = "black", fill = "#EB9063", padding = margin(5, 5, 5, 5)), 
        panel.background = element_rect(fill = "white", colour = "black"),
        plot.background = element_rect(fill = NA, colour = NA))+
  ggtitle("2-stroke") +
  scale_fill_manual(values = c("Pre" = "darkgrey", "During" = "#E25A00", "Post" = "#F2B389")) +
  scale_color_manual(values = c("Pre" = "#242424", "During" = "#542709", "Post" = "#66422a"))

shrimp_out_fig_two
```
####2.4.3.6 Combining plots into one figure
```{r}
shrimp_out_plot3 <- shrimp_out_fig_control + shrimp_out_fig_four + shrimp_out_fig_two + 
  plot_layout(ncol = 3)

shrimp_out_plot3
ggsave("./3-figs/shrimp_out_plot3.png", plot = shrimp_out_plot3, units = "px", width = 2200, height = 1400, dpi = 300)
```

###2.4.4 Effects of covariates on the proportion of time shrimp spent outside of the burrow
####2.4.4.1 Goby number
```{r}
shrimp_out_goby_number_contrasts <- shrimp_out_model |>
  emmeans(~ goby_number, 
          epred = TRUE,
          re_formla = NULL,
          rg.limit =14000) |> 
  contrast(method = "revpairwise")

shrimp_out_goby_number_contrasts
```

####2.4.4.2 Shrimp number
```{r}
shrimp_out_shrimp_number_contrasts <- shrimp_out_model |>
  emmeans(~ shrimp_number, 
          epred = TRUE,
          re_formla = NULL,
          rg.limit =14000) |> 
  contrast(method = "revpairwise")

shrimp_out_shrimp_number_contrasts

shrimp_out_shrimp_number_estimates <- shrimp_out_model |>
  emmeans(~ shrimp_number, 
          epred = TRUE,
          re_formla = NULL,
          rg.limit =14000)

shrimp_out_shrimp_number_estimates

shrimp_out_shrimp_number_contrasts <- shrimp_out_model |>
  emmeans(~ shrimp_number, 
          epred = TRUE,
          re_formla = NULL,
          rg.limit =14000) |> 
  contrast(method = "revpairwise") |> 
  gather_emmeans_draws()

ggplot(data = shrimp_out_shrimp_number_contrasts, aes(x = contrast, y = .value )) +
  stat_halfeye()
```
####2.4.4.3 Shrimp species
```{r}
shrimp_out_shrimp_species_contrasts <- shrimp_out_model |>
  emmeans(~ shrimp_species, 
          epred = TRUE,
          re_formla = NULL,
          rg.limit =14000) |> 
  contrast(method = "revpairwise")

shrimp_out_shrimp_species_contrasts

shrimp_out_shrimp_species_contrasts <- shrimp_out_model |>
  emmeans(~ shrimp_species, 
          epred = TRUE,
          re_formla = NULL,
          rg.limit =14000) |> 
  contrast(method = "revpairwise") |> 
  gather_emmeans_draws()

ggplot(data = shrimp_out_shrimp_species_contrasts, aes(x = contrast, y = .value )) +
  stat_halfeye()
```

##2.5 Modelling the proportion of time shrimp spent in contact with the goby
###2.5.1 Checking the response variable distribution
```{r}
shrimp_contact_df <- final_df |> 
  mutate(is_zero = shrimp_contact == 0)

shrimp_contact_plot <- shrimp_contact_df |> 
  mutate(shrimp_contact = ifelse(is_zero, -0.1, shrimp_contact)) |> 
  ggplot(aes(x = shrimp_contact)) +
  geom_histogram(aes(fill = is_zero), binwidth = 0.05, 
                 boundary = 0, color = "white") +
  geom_vline(xintercept = 0)
shrimp_contact_plot
# Distribution is a zero-one inflated beta distribution, because the value for shrimp_contact is bounded by 0-1 as it is a proportion of the total time and there is values of 0 and 1.
```

###2.5.2 Running the model
```{r}
shrimp_contact_model <- brm(
  bf(shrimp_contact ~ stroke*stage + shrimp_number + goby_number + shrimp_species + (1|site/burrow),
     phi ~ stroke*stage + shrimp_number + goby_number + shrimp_species + (1|site/burrow),
     zoi ~ stroke*stage + shrimp_number + goby_number + shrimp_species + (1|site/burrow),
     coi ~ stroke*stage + shrimp_number + goby_number + shrimp_species + (1|site/burrow)
     ),
  data = final_df,
  prior = priors,
  family = zero_one_inflated_beta(),
  control = list(adapt_delta = 0.99, max_treedepth = 12),
  chains = 4, iter = 6000, warmup = 1000, seed = 12345, backend = "cmdstanr", cores = 4, threads = threading(2),
  file = "shrimp_contact_model"
)

summary(shrimp_contact_model)
```

###2.5.3 Model fit and diagnostics
####2.5.3.1 Model convergence
```{r}
extract_rhats(shrimp_contact_model)
```

####2.5.3.2 Model fit
```{r}
#posterior predictive checks
pp_check(shrimp_contact_model)

pp_check(shrimp_contact_model, type='stat', stat='mean')

#fitted vs predicted values
summary(fitted(shrimp_contact_model))
summary(predict(shrimp_contact_model))
```

###2.5.2 Posterior marginal effects
####2.5.2.1 Pairwise contrasts of stroke and stage
```{r}
shrimp_contact_contrasts <- shrimp_contact_model |>
  emmeans(~ stroke*stage, 
          epred = TRUE,
          re_formla = NULL,
          rg.limit =14000) |> 
  contrast(method = "revpairwise")

# Extract the relevant contrast information
shrimp_contact_contrasts <- as.data.frame(shrimp_contact_contrasts) |> 
  dplyr::filter(contrast %in% c("control Pre - 2 Pre", "control Pre - 4 Pre","4 Pre - 2 Pre", "control Pre - control During", "control Pre - control Post", "4 Pre - 4 During", "4 Pre - 4 Post", "2 Pre - 2 During", "2 Pre - 2 Post"))

shrimp_contact_contrasts

shrimp_contact_contrasts_draws <- shrimp_contact_model  |>
  emmeans(~ stroke*stage, 
          epred = TRUE,
          re_formula = NA) |> 
  contrast(method = "revpairwise") |> 
  gather_emmeans_draws()

shrimp_contact_temporal_contrasts <- shrimp_contact_contrasts_draws |> 
  dplyr::filter(contrast %in% c("control Pre - control During", "control Pre - control Post","control Post - control During" , "4 Pre - 4 During", "4 Pre - 4 Post", "4 Post - 4 During", "2 Pre - 2 During", "2 Pre - 2 Post", "2 Post - 2 During")) |> 
  mutate(stroke = ifelse(contrast %in% c("control Pre - control During", "control Pre - control Post","control Post - control During"), "Control", ifelse(contrast %in% c("4 Pre - 4 During", "4 Pre - 4 Post", "4 Post - 4 During"), "4", "2")))

shrimp_contact_temporal_contrasts$stroke <- factor(shrimp_contact_temporal_contrasts$stroke, levels = c("Control", "4", "2"))

shrimp_contact_temporal_fig<- ggplot(shrimp_contact_temporal_contrasts, aes(x = .value, y = factor(contrast, levels = c("control Pre - control During", "control Pre - control Post","control Post - control During" , "4 Pre - 4 During", "4 Pre - 4 Post", "4 Post - 4 During", "2 Pre - 2 During", "2 Pre - 2 Post", "2 Post - 2 During")), color = stroke, fill = stroke)) +
   geom_vline(xintercept = 0, color = "black", size = 0.5) +
  stat_halfeye(.width = c(0.89, 0.95)) +
  scale_color_manual(values = c(colour11, colour12, colour13)) +
  scale_fill_manual(values = alpha(c(colour1, colour2, colour3), 0.8)) +
  labs(x = "Difference in the proportion of time spent out of the burrow",
       y = "Contrasts of the time periods relative to noise exposure",
       fill = "Noise treatment", color = "Noise treatment",
       fill_ramp = "Credible interval") +
  coord_flip() +
  theme(legend.position = "top", 
        axis.text.x = element_text(angle = 50, hjust =1)) +
  scale_y_discrete(labels = c("Pre - During", "Pre - Post","Post - During" , "Pre - During", "Pre - Post", "Post - During", "Pre - During", "Pre - Post", "Post - During")) +
  scale_x_continuous(breaks = seq(-0.2, 0.3, 0.1), labels = c("-0.2","-0.1" , "0", "0.1", "0.2", "0.3"), limits = c(-0.2, 0.3))

shrimp_contact_temporal_fig
```

###2.5.3 Extracting and plotting model estimates
####2.5.3.1 Extracting estimates
```{r}
shrimp_contact_model_estimates <- shrimp_contact_model |> 
  emmeans(~ stroke*stage,
          epred = TRUE,
          rg.limit = 14000,
          re_formula = NA)

shrimp_contact_model_estimates <- as.data.frame(shrimp_contact_model_estimates)

shrimp_contact_model_estimates

write.csv(shrimp_contact_model_estimates, "./2-data/shrimp_contact_model_estimates.csv")

#Storing the model estimate means for each group to plot later on
shrimp_contact_model_estimates_control <- shrimp_contact_model_estimates |> 
  dplyr::filter(stroke == "control") |> 
  rename(.value = emmean)

shrimp_contact_model_estimates_four <- shrimp_contact_model_estimates |> 
  dplyr::filter(stroke == "4")|> 
  rename(.value = emmean)

shrimp_contact_model_estimates_two <- shrimp_contact_model_estimates |> 
  dplyr::filter(stroke == "2")|> 
  rename(.value = emmean)
```
####2.5.3.2 Plotting model estimates and raw data
```{r}
shrimp_contact_model_draws <- shrimp_contact_model |> 
  emmeans(~ stroke*stage,
          epred = TRUE,
          rg.limit = 14000,
          re_formula = NA) |> 
  gather_emmeans_draws()

shrimp_contact_model_draws <- shrimp_contact_model_draws |> 
  rename(emmean = .value)

shrimp_contact_raw_data <- final_df |> 
  rename(emmean = shrimp_contact)

shrimp_contact_raw_data$stage <- factor(shrimp_contact_raw_data$stage, levels = c("Pre", "During", "Post"))
shrimp_contact_raw_data$stroke <- factor(shrimp_contact_raw_data$stroke, levels = c("control", "4", "2"))

shrimp_contact_model_estimates$stage <- factor(shrimp_contact_model_estimates$stage, levels = c("Pre", "During", "Post"))
shrimp_contact_model_estimates$stroke <- factor(shrimp_contact_model_estimates$stroke, levels = c("control", "4", "2"))

shrimp_contact_model_draws$stage <- factor(shrimp_contact_model_draws$stage, levels = c("Pre", "During", "Post"))
shrimp_contact_model_draws$stroke <- factor(shrimp_contact_model_draws$stroke, levels = c("control", "4", "2"))

shrimp_contact_means <- shrimp_contact_raw_data |> 
  group_by(stage, stroke) |> 
  summarise(emmean = mean(emmean))

shrimp_contact_fig_together <- ggplot(data = shrimp_contact_raw_data, aes(x = stage, y = emmean, color = stroke, fill = stroke)) +
  geom_point(data = shrimp_contact_means, alpha = 1, position = position_dodgenudge(width = 0.8, x = -0.09), size = 9, shape = "_") +
  geom_half_point(alpha = 0.3, position = position_dodgenudge(width = 0.8, x = -0.03), side = "l", range_scale = 0.5) +
  geom_half_violin(data = shrimp_contact_model_draws, aes(x = stage, y = emmean, color = stroke, fill = stroke), alpha = 0.6, position = position_dodge(width = 0.8), side = "r", size = 0) +
  geom_point(data = shrimp_contact_model_estimates, aes(x = stage, y = emmean, color = stroke), position = position_dodge(width = 0.8), size = 3) +
  geom_errorbar(data = shrimp_contact_model_estimates, aes(x = stage, y = emmean, ymin = lower.HPD, ymax = upper.HPD, color = stroke), position = position_dodge(width = 0.8), width = 0, size = 1.25) +
  scale_fill_manual(values = c("control" = "#76A34A", "4" = "#F29E00", "2" = "#E25A00"), name = "Noise exposure", guide = guide_legend(override.aes = list(size = 3))) +
  scale_color_manual(values = c("control" = "#2A330E", "4" = "#805213", "2" = "#542709"), name = "Noise exposure", guide = guide_legend(override.aes = list(size = 3))) +
  scale_x_discrete(labels = c("Pre", "During", "Post")) +
  labs(title = "The effects of boat noise on goby-shrimp contact",
       y = "Proportion of time the in contact",
       x = "Experimental phase relative to noise exposure") +
  theme_classic() +
  theme(
    axis.text.x = element_text(size = 10, margin = margin(t = 5)),  
    axis.text.y = element_text(size = 10, margin = margin(r = 5)),  
    axis.title.x = element_text(size = 13, margin = margin(t = 10)),
    axis.title.y = element_text(size = 13, margin = margin(r = 10)),
    legend.text = element_text(size = 10),
    legend.title = element_text(size = 10),
    legend.position = "right",
    legend.justification = "top",
    plot.title = element_text(hjust = 0.5, size = 16),  
    plot.margin = margin(10, 10, 10, 10)
  )

shrimp_contact_fig_together
ggsave("./3-figs/shrimp_contact_fig_together.png", plot = shrimp_contact_fig_together, units = "px", width = 2200, height = 1400, dpi = 300)
```

####2.5.3.3 Plotting control estimates and raw data
```{r}
shrimp_contact_model_draws <- shrimp_contact_model |> 
  emmeans(~ stroke*stage,
          epred = TRUE,
          rg.limit = 14000,
          re_formula = NA) |> 
  gather_emmeans_draws()

shrimp_contact_raw_data <- final_df |>
  rename(.value = shrimp_contact)

shrimp_contact_model_draws$stage <- factor(shrimp_contact_model_draws$stage, levels = c("Pre", "During", "Post"))
shrimp_contact_model_draws$stroke <- factor(shrimp_contact_model_draws$stroke, levels = c("control", "4", "2"))

shrimp_contact_control_model_draws <- shrimp_contact_model_draws |> 
  dplyr::filter(stroke == "control")

goby_control_raw_data <- final_df |> 
  dplyr::filter(stroke =="control") |> 
  rename(.value = shrimp_contact)

goby_control_means <- goby_control_raw_data |> 
  group_by(stage) |> 
  summarise(.value = mean(.value))

goby_control_raw_data$stage <- factor(goby_control_raw_data$stage, levels = c("Pre", "During", "Post"))

#assigning a value for the baseline mean
shrimp_contact_model_estimates_control_intercept <- shrimp_contact_model_estimates_control %>%
  dplyr::filter(stage == "Pre") %>%
  pull(.value)

shrimp_contact_fig_control <- ggplot(goby_control_raw_data, aes(y = .value, x = stage, fill = stage, color = stage)) +
  geom_half_violin(data = shrimp_contact_control_model_draws, alpha = 0.7, size = 0, side = "r", position = position_nudge(x = 0.2)) +
  geom_jitter(position = position_jitternudge(x = -0.1, width = 0.1, nudge.from = "jittered"), alpha = 0.4, size = 2) +
  geom_crossbar(alpha = 0.6, size = 0.8, data = shrimp_contact_model_estimates_control, aes(ymin = lower.HPD, ymax = upper.HPD), width = 0, position = position_nudge(x = 0.2))+
  geom_point(pch = 23, stroke = 1.2, data = shrimp_contact_model_estimates_control, size = 2, position = position_nudge(x = 0.2)) +
  geom_point(data = goby_control_means, position = position_dodgenudge(width = 0.8, x = -0.1), size = 9, shape = "_") +
  labs(x="Period relative to treatment",
       y="Proportion of time in contact") +
  theme(legend.position = "none", 
        axis.title.x = element_blank(), 
        axis.text.x = element_text(size = 13),  
        axis.text.y = element_text(size = 13),
        axis.title.y = element_text(size = 16, face = "bold", margin = margin(t = 0, b = 0, l = 0, r = 10)),
        plot.title = element_textbox_simple(face = "bold", size = 16, halign = 0.5, linewidth = 0.3, linetype = 1, color = "black", fill = "#A3C088", padding = margin(5, 5, 5, 5)), 
        panel.background = element_rect(fill = "white", colour = "black"),
        plot.background = element_rect(fill = NA, colour = NA))+
  ggtitle("Control")+
  scale_fill_manual(values = c("Pre" = "darkgrey", "During" = "#76A34A", "Post" = "#D3D9A7")) +
  scale_color_manual(values = c("Pre" = "#242424", "During" = "#2A330E", "Post" = "#5d634c"))

shrimp_contact_fig_control
```
####2.5.3.4 Plotting 4-stroke estimates and raw data
```{r}
shrimp_contact_control_model_draws <- shrimp_contact_model_draws |> 
  dplyr::filter(stroke == "4")

goby_four_raw_data <- final_df |> 
  dplyr::filter(stroke =="4") |> 
  rename(.value = shrimp_contact)

goby_four_raw_data$stage <- factor(goby_four_raw_data$stage, levels = c("Pre", "During", "Post"))

goby_four_means <- goby_four_raw_data |> 
  group_by(stage) |> 
  summarise(.value = mean(.value))

shrimp_contact_fig_four <- ggplot(goby_four_raw_data, aes(y = .value, x = stage, fill = stage, color = stage)) +
  geom_half_violin(data = shrimp_contact_control_model_draws, alpha = 0.7, size = 0, side = "r", position = position_nudge(x = 0.2)) +
  geom_jitter(position = position_jitternudge(x = -0.1, width = 0.1, nudge.from = "jittered"), alpha = 0.4, size = 2) +
  geom_crossbar(alpha = 0.6, size = 0.8, data = shrimp_contact_model_estimates_four, aes(ymin = lower.HPD, ymax = upper.HPD), width = 0, position = position_nudge(x = 0.2))+
  geom_point(pch = 23, stroke = 1.2, data = shrimp_contact_model_estimates_four, size = 2, position = position_nudge(x = 0.2)) +
  geom_point(data = goby_four_means, position = position_dodgenudge(width = 0.8, x = -0.1), size = 9, shape = "_") +
  xlab("Experimental phase relative to noise exposure")+
  theme(legend.position = "none", 
        axis.title.y = element_blank(),  
        axis.text.x = element_text(size = 13),  
        axis.text.y = element_blank(), 
        axis.ticks.y = element_blank(), 
        axis.title.x = element_text(size = 16, face = "bold", margin = margin(t = 15, b = 0, l = 0, r = 0)),
        plot.title = element_textbox_simple(face = "bold", size = 16, halign = 0.5, linewidth = 0.3, linetype = 1, color = "black", fill = "#F6BC65", padding = margin(5, 5, 5, 5)), 
        panel.background = element_rect(fill = "white", colour = "black"),
        plot.background = element_rect(fill = NA, colour = NA))+
  ggtitle("4-stroke")+
  scale_fill_manual(values = c("Pre" = "darkgrey", "During" = "#F29E00", "Post" = "#F2D5A0")) +
  scale_color_manual(values = c("Pre" = "#242424", "During" = "#805213", "Post" = "#70532b"))

shrimp_contact_fig_four
```

####2.5.3.5 Plotting 2-stroke estimates and raw data
```{r}
shrimp_contact_control_model_draws <- shrimp_contact_model_draws |> 
  dplyr::filter(stroke == "2")

goby_two_raw_data <- final_df |> 
  dplyr::filter(stroke =="2") |> 
  rename(.value = shrimp_contact)

goby_two_raw_data$stage <- factor(goby_two_raw_data$stage, levels = c("Pre", "During", "Post"))

goby_two_means <- goby_two_raw_data |> 
  group_by(stage) |> 
  summarise(.value = mean(.value))

shrimp_contact_fig_two <- ggplot(goby_two_raw_data, aes(y = .value, x = stage, fill = stage, color = stage)) +
  geom_half_violin(data = shrimp_contact_control_model_draws, alpha = 0.7, size = 0, side = "r", position = position_nudge(x = 0.2)) +
  geom_jitter(position = position_jitternudge(x = -0.1, width = 0.1, nudge.from = "jittered"), alpha = 0.4, size = 2) +
  geom_crossbar(alpha = 0.6, size = 0.8, data = shrimp_contact_model_estimates_two, aes(ymin = lower.HPD, ymax = upper.HPD), width = 0, position = position_nudge(x = 0.2))+
  geom_point(pch = 23, stroke = 1.2, data = shrimp_contact_model_estimates_two, size = 2, position = position_nudge(x = 0.2)) +
  geom_point(data = goby_two_means, position = position_dodgenudge(width = 0.8, x = -0.1), size = 9, shape = "_") +
  theme(legend.position = "none", 
        axis.title.x = element_blank(),  
        axis.text.x = element_text(size = 13),
        axis.title.y = element_blank(), 
        axis.text.y = element_blank(), 
        axis.ticks.y = element_blank(), 
        plot.title = element_textbox_simple(face = "bold", size = 16, halign = 0.5, linewidth = 0.3, linetype = 1, color = "black", fill = "#EB9063", padding = margin(5, 5, 5, 5)), 
        panel.background = element_rect(fill = "white", colour = "black"),
        plot.background = element_rect(fill = NA, colour = NA))+
  ggtitle("2-stroke") +
  scale_fill_manual(values = c("Pre" = "darkgrey", "During" = "#E25A00", "Post" = "#F2B389")) +
  scale_color_manual(values = c("Pre" = "#242424", "During" = "#542709", "Post" = "#66422a"))

shrimp_contact_fig_two
```
####2.5.3.6 Combining plots into one figure
```{r}
shrimp_contact_plot3 <- shrimp_contact_fig_control + shrimp_contact_fig_four + shrimp_contact_fig_two + 
  plot_layout(ncol = 3)

shrimp_contact_plot3
ggsave("./3-figs/shrimp_contact_plot3.png", plot = shrimp_contact_plot3, units = "px", width = 2200, height = 1400, dpi = 300)
```

###2.5.4 Modelling the effect of covariates on the proportion of time spent in contact
####2.5.4.1 Goby number
```{r}
shrimp_contact_goby_number_contrasts <- shrimp_contact_model |>
  emmeans(~ goby_number, 
          epred = TRUE,
          re_formla = NULL,
          rg.limit =14000) |> 
  contrast(method = "revpairwise")

shrimp_contact_goby_number_contrasts
```

####2.5.4.2 Shrimp number
```{r}
shrimp_contact_shrimp_number_contrasts <- shrimp_contact_model |>
  emmeans(~ shrimp_number, 
          epred = TRUE,
          re_formla = NULL,
          rg.limit =14000) |> 
  contrast(method = "revpairwise")

shrimp_contact_shrimp_number_contrasts
```

####2.5.4.2 Shrimp species
```{r}
shrimp_contact_shrimp_species_contrasts <- shrimp_contact_model |>
  emmeans(~ shrimp_species, 
          epred = TRUE,
          re_formla = NULL,
          rg.limit =14000) |> 
  contrast(method = "revpairwise")

shrimp_contact_shrimp_species_contrasts
```

#Alternate graphs
## Goby out
```{r}
goby_out_model_draws <- goby_out_model |> 
  emmeans(~ stroke*stage,
          epred = TRUE,
          rg.limit = 14000,
          re_formula = NA) |> 
  gather_emmeans_draws()

goby_out_raw_data <- final_df |>
  rename(.value = goby_out)

goby_out_model_draws$stage <- factor(goby_out_model_draws$stage, levels = c("Pre", "During", "Post"))
goby_out_model_draws$stroke <- factor(goby_out_model_draws$stroke, levels = c("control", "4", "2"))

goby_out_control_model_draws <- goby_out_model_draws |> 
  dplyr::filter(stroke == "control")

goby_control_raw_data <- final_df |> 
  dplyr::filter(stroke =="control") |> 
  rename(.value = goby_out)

goby_control_means <- goby_control_raw_data |> 
  group_by(stage) |> 
  summarise(.value = mean(.value))

goby_control_raw_data$stage <- factor(goby_control_raw_data$stage, levels = c("Pre", "During", "Post"))

#assigning a value for the baseline mean
goby_out_model_estimates_control_intercept <- goby_out_model_estimates_control %>%
  dplyr::filter(stage == "Pre") %>%
  pull(.value)

goby_out_fig_control2 <- ggplot(goby_control_raw_data, aes(y = .value, x = stage, fill = stage, color = stage)) +
  geom_jitter(position = position_jitternudge(x = -0.1, width = 0.1, nudge.from = "jittered"), alpha = 0.4, size = 2) +
  geom_crossbar(alpha = 0.6, size = 0.8, data = goby_out_model_estimates_control, aes(ymin = lower.HPD, ymax = upper.HPD), width = 0, position = position_nudge(x = 0.2))+
  geom_point(pch = 23, stroke = 1.2, data = goby_out_model_estimates_control, size = 2, position = position_nudge(x = 0.2)) +
  geom_point(data = goby_control_means, position = position_dodgenudge(width = 0.8, x = -0.1), size = 9, shape = "_") +
  labs(x="Period relative to treatment",
       y="Proportion of time out of burrow") +
  theme(legend.position = "none", 
        axis.title.x = element_blank(), 
        axis.text.x = element_text(size = 13),  
        axis.text.y = element_text(size = 13),
        axis.title.y = element_text(size = 16, face = "bold", margin = margin(t = 0, b = 0, l = 0, r = 10)),
        plot.title = element_textbox_simple(face = "bold", size = 16, halign = 0.5, linewidth = 0.3, linetype = 1, color = "black", fill = "#A3C088", padding = margin(5, 5, 5, 5)), 
        panel.background = element_rect(fill = "white", colour = "black"),
        plot.background = element_rect(fill = NA, colour = NA))+
  ggtitle("Control")+
  scale_fill_manual(values = c("Pre" = "darkgrey", "During" = "#76A34A", "Post" = "#D3D9A7")) +
  scale_color_manual(values = c("Pre" = "#242424", "During" = "#2A330E", "Post" = "#5d634c"))

goby_out_fig_control

goby_out_control_model_draws <- goby_out_model_draws |> 
  dplyr::filter(stroke == "4")

goby_four_raw_data <- final_df |> 
  dplyr::filter(stroke =="4") |> 
  rename(.value = goby_out)

goby_four_raw_data$stage <- factor(goby_four_raw_data$stage, levels = c("Pre", "During", "Post"))

goby_four_means <- goby_four_raw_data |> 
  group_by(stage) |> 
  summarise(.value = mean(.value))

goby_out_fig_four2 <- ggplot(goby_four_raw_data, aes(y = .value, x = stage, fill = stage, color = stage)) +
  geom_jitter(position = position_jitternudge(x = -0.1, width = 0.1, nudge.from = "jittered"), alpha = 0.4, size = 2) +
  geom_crossbar(alpha = 0.6, size = 0.8, data = goby_out_model_estimates_four, aes(ymin = lower.HPD, ymax = upper.HPD), width = 0, position = position_nudge(x = 0.2))+
  geom_point(pch = 23, stroke = 1.2, data = goby_out_model_estimates_four, size = 2, position = position_nudge(x = 0.2)) +
  geom_point(data = goby_four_means, position = position_dodgenudge(width = 0.8, x = -0.1), size = 9, shape = "_") +
  xlab("Experimental phase relative to noise exposure")+
  theme(legend.position = "none", 
        axis.title.y = element_blank(),  
        axis.text.x = element_text(size = 13),  
        axis.text.y = element_blank(), 
        axis.ticks.y = element_blank(), 
        axis.title.x = element_text(size = 16, face = "bold", margin = margin(t = 15, b = 0, l = 0, r = 0)),
        plot.title = element_textbox_simple(face = "bold", size = 16, halign = 0.5, linewidth = 0.3, linetype = 1, color = "black", fill = "#F6BC65", padding = margin(5, 5, 5, 5)), 
        panel.background = element_rect(fill = "white", colour = "black"),
        plot.background = element_rect(fill = NA, colour = NA))+
  ggtitle("4-stroke")+
  scale_fill_manual(values = c("Pre" = "darkgrey", "During" = "#F29E00", "Post" = "#F2D5A0")) +
  scale_color_manual(values = c("Pre" = "#242424", "During" = "#805213", "Post" = "#70532b"))

goby_out_fig_four

goby_out_control_model_draws <- goby_out_model_draws |> 
  dplyr::filter(stroke == "2")

goby_two_raw_data <- final_df |> 
  dplyr::filter(stroke =="2") |> 
  rename(.value = goby_out)

goby_two_raw_data$stage <- factor(goby_two_raw_data$stage, levels = c("Pre", "During", "Post"))

goby_two_means <- goby_two_raw_data |> 
  group_by(stage) |> 
  summarise(.value = mean(.value))

goby_out_fig_two2 <- ggplot(goby_two_raw_data, aes(y = .value, x = stage, fill = stage, color = stage)) +
  geom_jitter(position = position_jitternudge(x = -0.1, width = 0.1, nudge.from = "jittered"), alpha = 0.4, size = 2) +
  geom_crossbar(alpha = 0.6, size = 0.8, data = goby_out_model_estimates_two, aes(ymin = lower.HPD, ymax = upper.HPD), width = 0, position = position_nudge(x = 0.2))+
  geom_point(pch = 23, stroke = 1.2, data = goby_out_model_estimates_two, size = 2, position = position_nudge(x = 0.2)) +
  geom_point(data = goby_two_means, position = position_dodgenudge(width = 0.8, x = -0.1), size = 9, shape = "_") +
  theme(legend.position = "none", 
        axis.title.x = element_blank(),  
        axis.text.x = element_text(size = 13),
        axis.title.y = element_blank(), 
        axis.text.y = element_blank(), 
        axis.ticks.y = element_blank(), 
        plot.title = element_textbox_simple(face = "bold", size = 16, halign = 0.5, linewidth = 0.3, linetype = 1, color = "black", fill = "#EB9063", padding = margin(5, 5, 5, 5)), 
        panel.background = element_rect(fill = "white", colour = "black"),
        plot.background = element_rect(fill = NA, colour = NA))+
  ggtitle("2-stroke") +
  scale_fill_manual(values = c("Pre" = "darkgrey", "During" = "#E25A00", "Post" = "#F2B389")) +
  scale_color_manual(values = c("Pre" = "#242424", "During" = "#542709", "Post" = "#66422a"))

goby_out_fig_two

goby_out_plot2 <- goby_out_fig_control2 + goby_out_fig_four2 + goby_out_fig_two2 + 
  plot_layout(ncol = 3)

goby_out_plot2
ggsave("./3-figs/goby_out_plot2.png", plot = goby_out_plot2, units = "px", width = 2200, height = 1400, dpi = 300)
```

##Shrmip out
```{r}
shrimp_out_model_draws <- shrimp_out_model |> 
  emmeans(~ stroke*stage,
          epred = TRUE,
          rg.limit = 14000,
          re_formula = NA) |> 
  gather_emmeans_draws()

shrimp_out_raw_data <- final_df |>
  rename(.value = shrimp_out)

shrimp_out_model_draws$stage <- factor(shrimp_out_model_draws$stage, levels = c("Pre", "During", "Post"))
shrimp_out_model_draws$stroke <- factor(shrimp_out_model_draws$stroke, levels = c("control", "4", "2"))

shrimp_out_control_model_draws <- shrimp_out_model_draws |> 
  dplyr::filter(stroke == "control")

shrimp_control_raw_data <- final_df |> 
  dplyr::filter(stroke =="control") |> 
  rename(.value = shrimp_out)

shrimp_control_means <- shrimp_control_raw_data |> 
  group_by(stage) |> 
  summarise(.value = mean(.value))

shrimp_control_raw_data$stage <- factor(shrimp_control_raw_data$stage, levels = c("Pre", "During", "Post"))

#assigning a value for the baseline mean
shrimp_out_model_estimates_control_intercept <- shrimp_out_model_estimates_control %>%
  dplyr::filter(stage == "Pre") %>%
  pull(.value)

shrimp_out_fig_control2 <- ggplot(shrimp_control_raw_data, aes(y = .value, x = stage, fill = stage, color = stage)) +
  geom_jitter(position = position_jitternudge(x = -0.1, width = 0.1, nudge.from = "jittered"), alpha = 0.4, size = 2) +
  geom_crossbar(alpha = 0.6, size = 0.8, data = shrimp_out_model_estimates_control, aes(ymin = lower.HPD, ymax = upper.HPD), width = 0, position = position_nudge(x = 0.2))+
  geom_point(pch = 23, stroke = 1.2, data = shrimp_out_model_estimates_control, size = 2, position = position_nudge(x = 0.2)) +
  geom_point(data = shrimp_control_means, position = position_dodgenudge(width = 0.8, x = -0.1), size = 9, shape = "_") +
  labs(x="Period relative to treatment",
       y="Proportion of time out of burrow") +
  theme(legend.position = "none", 
        axis.title.x = element_blank(), 
        axis.text.x = element_text(size = 13),  
        axis.text.y = element_text(size = 13),
        axis.title.y = element_text(size = 16, face = "bold", margin = margin(t = 0, b = 0, l = 0, r = 10)),
        plot.title = element_textbox_simple(face = "bold", size = 16, halign = 0.5, linewidth = 0.3, linetype = 1, color = "black", fill = "#A3C088", padding = margin(5, 5, 5, 5)), 
        panel.background = element_rect(fill = "white", colour = "black"),
        plot.background = element_rect(fill = NA, colour = NA))+
  ggtitle("Control")+
  scale_fill_manual(values = c("Pre" = "darkgrey", "During" = "#76A34A", "Post" = "#D3D9A7")) +
  scale_color_manual(values = c("Pre" = "#242424", "During" = "#2A330E", "Post" = "#5d634c"))

shrimp_out_fig_control

shrimp_out_control_model_draws <- shrimp_out_model_draws |> 
  dplyr::filter(stroke == "4")

shrimp_four_raw_data <- final_df |> 
  dplyr::filter(stroke =="4") |> 
  rename(.value = shrimp_out)

shrimp_four_raw_data$stage <- factor(shrimp_four_raw_data$stage, levels = c("Pre", "During", "Post"))

shrimp_four_means <- shrimp_four_raw_data |> 
  group_by(stage) |> 
  summarise(.value = mean(.value))

shrimp_out_fig_four2 <- ggplot(shrimp_four_raw_data, aes(y = .value, x = stage, fill = stage, color = stage)) +
  geom_jitter(position = position_jitternudge(x = -0.1, width = 0.1, nudge.from = "jittered"), alpha = 0.4, size = 2) +
  geom_crossbar(alpha = 0.6, size = 0.8, data = shrimp_out_model_estimates_four, aes(ymin = lower.HPD, ymax = upper.HPD), width = 0, position = position_nudge(x = 0.2))+
  geom_point(pch = 23, stroke = 1.2, data = shrimp_out_model_estimates_four, size = 2, position = position_nudge(x = 0.2)) +
  geom_point(data = shrimp_four_means, position = position_dodgenudge(width = 0.8, x = -0.1), size = 9, shape = "_") +
  xlab("Experimental phase relative to noise exposure")+
  theme(legend.position = "none", 
        axis.title.y = element_blank(),  
        axis.text.x = element_text(size = 13),  
        axis.text.y = element_blank(), 
        axis.ticks.y = element_blank(), 
        axis.title.x = element_text(size = 16, face = "bold", margin = margin(t = 15, b = 0, l = 0, r = 0)),
        plot.title = element_textbox_simple(face = "bold", size = 16, halign = 0.5, linewidth = 0.3, linetype = 1, color = "black", fill = "#F6BC65", padding = margin(5, 5, 5, 5)), 
        panel.background = element_rect(fill = "white", colour = "black"),
        plot.background = element_rect(fill = NA, colour = NA))+
  ggtitle("4-stroke")+
  scale_fill_manual(values = c("Pre" = "darkgrey", "During" = "#F29E00", "Post" = "#F2D5A0")) +
  scale_color_manual(values = c("Pre" = "#242424", "During" = "#805213", "Post" = "#70532b"))

shrimp_out_fig_four

shrimp_out_control_model_draws <- shrimp_out_model_draws |> 
  dplyr::filter(stroke == "2")

shrimp_two_raw_data <- final_df |> 
  dplyr::filter(stroke =="2") |> 
  rename(.value = shrimp_out)

shrimp_two_raw_data$stage <- factor(shrimp_two_raw_data$stage, levels = c("Pre", "During", "Post"))

shrimp_two_means <- shrimp_two_raw_data |> 
  group_by(stage) |> 
  summarise(.value = mean(.value))

shrimp_out_fig_two2 <- ggplot(shrimp_two_raw_data, aes(y = .value, x = stage, fill = stage, color = stage)) +
  geom_jitter(position = position_jitternudge(x = -0.1, width = 0.1, nudge.from = "jittered"), alpha = 0.4, size = 2) +
  geom_crossbar(alpha = 0.6, size = 0.8, data = shrimp_out_model_estimates_two, aes(ymin = lower.HPD, ymax = upper.HPD), width = 0, position = position_nudge(x = 0.2))+
  geom_point(pch = 23, stroke = 1.2, data = shrimp_out_model_estimates_two, size = 2, position = position_nudge(x = 0.2)) +
  geom_point(data = shrimp_two_means, position = position_dodgenudge(width = 0.8, x = -0.1), size = 9, shape = "_") +
  theme(legend.position = "none", 
        axis.title.x = element_blank(),  
        axis.text.x = element_text(size = 13),
        axis.title.y = element_blank(), 
        axis.text.y = element_blank(), 
        axis.ticks.y = element_blank(), 
        plot.title = element_textbox_simple(face = "bold", size = 16, halign = 0.5, linewidth = 0.3, linetype = 1, color = "black", fill = "#EB9063", padding = margin(5, 5, 5, 5)), 
        panel.background = element_rect(fill = "white", colour = "black"),
        plot.background = element_rect(fill = NA, colour = NA))+
  ggtitle("2-stroke") +
  scale_fill_manual(values = c("Pre" = "darkgrey", "During" = "#E25A00", "Post" = "#F2B389")) +
  scale_color_manual(values = c("Pre" = "#242424", "During" = "#542709", "Post" = "#66422a"))

shrimp_out_fig_two

shrimp_out_plot2 <- shrimp_out_fig_control2 + shrimp_out_fig_four2 + shrimp_out_fig_two2 + 
  plot_layout(ncol = 3)

shrimp_out_plot2
ggsave("./3-figs/shrimp_out_plot2.png", plot = shrimp_out_plot2, units = "px", width = 2200, height = 1400, dpi = 300)
```

##Shrmip contact
```{r}
shrimp_contact_model_draws <- shrimp_contact_model |> 
  emmeans(~ stroke*stage,
          epred = TRUE,
          rg.limit = 14000,
          re_formula = NA) |> 
  gather_emmeans_draws()

shrimp_contact_raw_data <- final_df |>
  rename(.value = shrimp_contact)

shrimp_contact_model_draws$stage <- factor(shrimp_contact_model_draws$stage, levels = c("Pre", "During", "Post"))
shrimp_contact_model_draws$stroke <- factor(shrimp_contact_model_draws$stroke, levels = c("control", "4", "2"))

shrimp_contact_control_model_draws <- shrimp_contact_model_draws |> 
  dplyr::filter(stroke == "control")

shrimp_control_raw_data <- final_df |> 
  dplyr::filter(stroke =="control") |> 
  rename(.value = shrimp_contact)

shrimp_control_means <- shrimp_control_raw_data |> 
  group_by(stage) |> 
  summarise(.value = mean(.value))

shrimp_control_raw_data$stage <- factor(shrimp_control_raw_data$stage, levels = c("Pre", "During", "Post"))

#assigning a value for the baseline mean
shrimp_contact_model_estimates_control_intercept <- shrimp_contact_model_estimates_control %>%
  dplyr::filter(stage == "Pre") %>%
  pull(.value)

shrimp_contact_fig_control2 <- ggplot(shrimp_control_raw_data, aes(y = .value, x = stage, fill = stage, color = stage)) +
  geom_jitter(position = position_jitternudge(x = -0.1, width = 0.1, nudge.from = "jittered"), alpha = 0.4, size = 2) +
  geom_crossbar(alpha = 0.6, size = 0.8, data = shrimp_contact_model_estimates_control, aes(ymin = lower.HPD, ymax = upper.HPD), width = 0, position = position_nudge(x = 0.2))+
  geom_point(pch = 23, stroke = 1.2, data = shrimp_contact_model_estimates_control, size = 2, position = position_nudge(x = 0.2)) +
  geom_point(data = shrimp_control_means, position = position_dodgenudge(width = 0.8, x = -0.1), size = 9, shape = "_") +
  labs(x="Period relative to treatment",
       y="Proportion of time contact of burrow") +
  theme(legend.position = "none", 
        axis.title.x = element_blank(), 
        axis.text.x = element_text(size = 13),  
        axis.text.y = element_text(size = 13),
        axis.title.y = element_text(size = 16, face = "bold", margin = margin(t = 0, b = 0, l = 0, r = 10)),
        plot.title = element_textbox_simple(face = "bold", size = 16, halign = 0.5, linewidth = 0.3, linetype = 1, color = "black", fill = "#A3C088", padding = margin(5, 5, 5, 5)), 
        panel.background = element_rect(fill = "white", colour = "black"),
        plot.background = element_rect(fill = NA, colour = NA))+
  ggtitle("Control")+
  scale_fill_manual(values = c("Pre" = "darkgrey", "During" = "#76A34A", "Post" = "#D3D9A7")) +
  scale_color_manual(values = c("Pre" = "#242424", "During" = "#2A330E", "Post" = "#5d634c"))

shrimp_contact_control_model_draws <- shrimp_contact_model_draws |> 
  dplyr::filter(stroke == "4")

shrimp_four_raw_data <- final_df |> 
  dplyr::filter(stroke =="4") |> 
  rename(.value = shrimp_contact)

shrimp_four_raw_data$stage <- factor(shrimp_four_raw_data$stage, levels = c("Pre", "During", "Post"))

shrimp_four_means <- shrimp_four_raw_data |> 
  group_by(stage) |> 
  summarise(.value = mean(.value))

shrimp_contact_fig_four2 <- ggplot(shrimp_four_raw_data, aes(y = .value, x = stage, fill = stage, color = stage)) +
  geom_jitter(position = position_jitternudge(x = -0.1, width = 0.1, nudge.from = "jittered"), alpha = 0.4, size = 2) +
  geom_crossbar(alpha = 0.6, size = 0.8, data = shrimp_contact_model_estimates_four, aes(ymin = lower.HPD, ymax = upper.HPD), width = 0, position = position_nudge(x = 0.2))+
  geom_point(pch = 23, stroke = 1.2, data = shrimp_contact_model_estimates_four, size = 2, position = position_nudge(x = 0.2)) +
  geom_point(data = shrimp_four_means, position = position_dodgenudge(width = 0.8, x = -0.1), size = 9, shape = "_") +
  xlab("Experimental phase relative to noise exposure")+
  theme(legend.position = "none", 
        axis.title.y = element_blank(),  
        axis.text.x = element_text(size = 13),  
        axis.text.y = element_blank(), 
        axis.ticks.y = element_blank(), 
        axis.title.x = element_text(size = 16, face = "bold", margin = margin(t = 15, b = 0, l = 0, r = 0)),
        plot.title = element_textbox_simple(face = "bold", size = 16, halign = 0.5, linewidth = 0.3, linetype = 1, color = "black", fill = "#F6BC65", padding = margin(5, 5, 5, 5)), 
        panel.background = element_rect(fill = "white", colour = "black"),
        plot.background = element_rect(fill = NA, colour = NA))+
  ggtitle("4-stroke")+
  scale_fill_manual(values = c("Pre" = "darkgrey", "During" = "#F29E00", "Post" = "#F2D5A0")) +
  scale_color_manual(values = c("Pre" = "#242424", "During" = "#805213", "Post" = "#70532b"))

shrimp_contact_control_model_draws <- shrimp_contact_model_draws |> 
  dplyr::filter(stroke == "2")

shrimp_two_raw_data <- final_df |> 
  dplyr::filter(stroke =="2") |> 
  rename(.value = shrimp_contact)

shrimp_two_raw_data$stage <- factor(shrimp_two_raw_data$stage, levels = c("Pre", "During", "Post"))

shrimp_two_means <- shrimp_two_raw_data |> 
  group_by(stage) |> 
  summarise(.value = mean(.value))

shrimp_contact_fig_two2 <- ggplot(shrimp_two_raw_data, aes(y = .value, x = stage, fill = stage, color = stage)) +
  geom_jitter(position = position_jitternudge(x = -0.1, width = 0.1, nudge.from = "jittered"), alpha = 0.4, size = 2) +
  geom_crossbar(alpha = 0.6, size = 0.8, data = shrimp_contact_model_estimates_two, aes(ymin = lower.HPD, ymax = upper.HPD), width = 0, position = position_nudge(x = 0.2))+
  geom_point(pch = 23, stroke = 1.2, data = shrimp_contact_model_estimates_two, size = 2, position = position_nudge(x = 0.2)) +
  geom_point(data = shrimp_two_means, position = position_dodgenudge(width = 0.8, x = -0.1), size = 9, shape = "_") +
  theme(legend.position = "none", 
        axis.title.x = element_blank(),  
        axis.text.x = element_text(size = 13),
        axis.title.y = element_blank(), 
        axis.text.y = element_blank(), 
        axis.ticks.y = element_blank(), 
        plot.title = element_textbox_simple(face = "bold", size = 16, halign = 0.5, linewidth = 0.3, linetype = 1, color = "black", fill = "#EB9063", padding = margin(5, 5, 5, 5)), 
        panel.background = element_rect(fill = "white", colour = "black"),
        plot.background = element_rect(fill = NA, colour = NA))+
  ggtitle("2-stroke") +
  scale_fill_manual(values = c("Pre" = "darkgrey", "During" = "#E25A00", "Post" = "#F2B389")) +
  scale_color_manual(values = c("Pre" = "#242424", "During" = "#542709", "Post" = "#66422a"))

shrimp_contact_plot2 <- shrimp_contact_fig_control2 + shrimp_contact_fig_four2 + shrimp_contact_fig_two2 + 
  plot_layout(ncol = 3)

shrimp_contact_plot2
ggsave("./3-figs/shrimp_contact_plot2.png", plot = shrimp_contact_plot2, units = "px", width = 2200, height = 1400, dpi = 300)
```